---
title: "**Nitinat in-season update**"
date: 'Last update: `r Sys.Date()`'
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_collapsed: false 
    toc_depth: 5
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning=F, message=F)


# Load packages ---------------------------------
library(tidyverse)


# Helpers --------------------------------- <<<UPDATE>>>
escapement_year <- 2024
options(scipen=9999)


# ================== COUNT DATA ================== 
# SIL data ---------------------------------
 nitinat.SIL <- readxl::read_excel(path=paste0("//ENT.dfo-mpo.ca/DFO-MPO/GROUP/PAC/PBS/Operations/SCA/SCD_Stad/WCVI/ESCAPEMENT/Data/", 
                                               escapement_year, 
                                               "/SILs/Area 21-22/Nitinat River/A22_Nitinat_River_", 
                                               escapement_year, 
                                               ".xlsx"),
                                   sheet="Nitinat River raw SIL entry", guess_max=1000) %>% 
   mutate_if(is.numeric, ~replace_na(., 0)) %>% 
   mutate_at(c("adults # holding", "adults live / habitat seen (# Est Live)"), ~round(.,0)) %>% 
   mutate(date = lubridate::ymd(date)) %>%
   print()


# ================== ENVIRO DATA ================== 
# Water data ---------------------------------
nitinat.water <- 
  # lapply(list.files(path = here::here("data", "enviro"),
  #                   pattern = "^NitinatWaterLevelHydrometData.*\\.xlsx$",
  #                   full.names = TRUE), 
  #        readxl::read_excel, sheet="Exported Data") %>% 
  #do.call("cbind",.) %>%
  lapply(list.files(path = here::here("data", "enviro"),
                    pattern = "NitinatWaterLevelHydrometData_20*",
                    full.names = TRUE),
         read.csv, skip=1) %>%
  do.call("rbind",.) %>%
  separate(Received, sep=" ", c("date", "time", "stamp")) %>%
  mutate(date2 = case_when(is.na(TL.Battery.V) ~ lubridate::mdy(date),
                          !is.na(TL.Battery.V) ~ lubridate::dmy(date)),
        year = lubridate::year(date2),
        month = lubridate::month(date2, label=T, abbr=T),
        DOY = lubridate::yday(date2),
        Water.Level.m = case_when(grepl("\\*", Water.Level.m) ~ NA,
                                  TRUE ~ as.numeric(Water.Level.m)),
        label_group = case_when(year<escapement_year ~ "Historical",
                                        year==escapement_year ~ as.character(year)),
        hour = as.numeric(stringr::str_sub(stringr::str_pad(time, width=5, pad = "0"), 1,2)),
        hour = case_when(hour==12 & stamp=="AM" ~ 0,
                         TRUE ~ hour)) %>%
  print()



# San Juan water data for comparison---
SJ.water <- full_join(
  read.csv(file=list.files(path = here::here("data", "enviro"),
                           pattern = "SANJUAN_historical*",
                           full.names = TRUE),
           skip=1) %>%
    mutate(PARAM = case_when(PARAM==1 ~ "discharge (cms)",
                             PARAM==2 ~ "level (m)"),
           Date = lubridate::ymd(Date),
           year = lubridate::year(Date),
           month = lubridate::month(Date, label=T, abbr=T),
           DOY = lubridate::yday(Date)),
  
    read.csv(file=list.files(path = here::here("data", "enviro"),
                           pattern = "SANJUAN_realtime*",
                           full.names = TRUE),
           skip=9) %>%
    mutate(Parameter = case_when(Parameter==46 ~ "level (m)"),
           Date = lubridate::ymd(stringr::str_sub(string=Date..PST., start=1, end=10)),
           time = stringr::str_sub(string=Date..PST., start=12, end=19),
           year = lubridate::year(Date..PST.),
           month = lubridate::month(Date..PST., label=T, abbr=T),
           DOY = lubridate::yday(Date..PST.)) %>% 
    rename(PARAM=Parameter,
           Value=Value..m.)
)



# ================== HISTORICAL DATA ================== 
# Run timing ---------------------------------
nit.rtStWk <- readxl::read_excel(path=#paste0(here::here("data", "escapement"), sep="/", 
                                      #       list.files(path=here::here("data", "escapement"),
                                      #                  pattern="run timing tables (updated)*")), 
                                   "//ENT.dfo-mpo.ca/DFO-MPO/GROUP/PAC/PBS/Operations/SCA/SCD_Stad/WCVI/ESCAPEMENT/Data/run timing tables (updated).xlsx",
                                 sheet="Nitinat", guess_max=10000, trim_ws=T, 
                                 range=cellranger::cell_limits(c(4, 1), c(NA, NA))) %>% 
  select(-c(`...24`:`...75`)) %>%
  rename(year=`...1`) %>% 
  filter(if_any(everything(), ~ !is.na(.)), !grepl("Table", year), !grepl("Year", year), !is.na(year)) %>% 
  mutate(species = c(rep("chinook", nrow(.)/3), rep("coho", nrow(.)/3), rep("chum", nrow(.)/3)),
         `92` = case_when(`92`=="1000-2000" ~ as.numeric(1500),
                          TRUE ~ as.numeric(`92`)),
         across(c(`81`:`124`), ~as.numeric(.)),
         plot_group = case_when(year==escapement_year ~ year,
                                year!=escapement_year ~ paste0(min(year), 
                                                               sep="-", (escapement_year-1)))) %>%
  pivot_longer(cols = c(`81`:`124`), names_to = "stat_week", values_to = "count") %>% 
  print()

nit.rtStWk$stat_week <- factor(nit.rtStWk$stat_week,
                               levels=c("81","82","83","84",
                                        "91","92","93","94",
                                        "101","102","103","104","105",
                                        "111","112","113","114","115",
                                        "121","122","123","124", ordered=T)) 
```

<br>


<br>

# **Water level and temperature**

The following water data are compiled from the Nitinat Hydromet operated by South Coast Stock Assessment: http://pacfish.ca/wcviweather/Content%20Pages/Nitinat/WaterLevel.aspx

As of `r Sys.Date()` the hydromet station has been inoperable since July 16, 2024. Work is underway to restore reporting before swims begin. For now, the following information is summarized from the backend database. Contact Katie with questions. 

```{r fig.cap='Nitinat hydromet water level (m). Historical shaded area represents daily maximum and minimum recorded values over the time series, while the line represents the mean water level. Current year line represents daily mean water level (blue). Water level is recorded hourly at the Nitinat River Hatchery pumphouse.', fig.width=8, fig.height=6}
ggplot()+
  geom_ribbon(data=nitinat.water %>% 
                filter(month %in% c("Aug","Sep","Oct", "Nov") & year!=escapement_year & Water.Level.m>0.1) %>%
                group_by(DOY) %>% 
                summarize(max_lvl=max(Water.Level.m,na.rm=T), min_lvl=min(Water.Level.m,na.rm=T)),
              aes(x=as.Date(DOY, origin="2021-12-31"), ymin=min_lvl, ymax=max_lvl), fill="gray80") +
  geomtextpath:: geom_textline(data=nitinat.water %>% 
                                filter(month %in% c("Aug","Sep","Oct", "Nov") & year!=escapement_year & Water.Level.m>0.1) %>%
                                group_by(DOY) %>% 
                                summarize(mean_lvl_hist=mean(Water.Level.m,na.rm=T)) %>%
                                mutate(label = paste0(min(nitinat.water$year), sep="-", escapement_year-1)),
                              aes(x=as.Date(DOY, origin="2021-12-31"), y=mean_lvl_hist, label=as.factor(label)), 
                              colour="gray40", linewidth=1, show.legend=F, hjust=0.92, text_smoothing=30) +
  geomtextpath:: geom_labelline(data=nitinat.water %>% 
                                filter(month %in% c("Aug","Sep","Oct", "Nov") & year==escapement_year & Water.Level.m>0.1) %>%
                                group_by(DOY) %>% 
                                summarize(mean_lvl=mean(Water.Level.m,na.rm=T)) %>% 
                                mutate(label=escapement_year),
                              aes(x=as.Date(DOY, origin="2021-12-31"), y=mean_lvl, label=as.factor(label)), 
                              colour="dodger blue", linewidth=1, hjust=0.02, show.legend=F) +
  labs(y="Water level (m)") +
  scale_colour_manual(values=c("dodger blue", "gray70")) +
  scale_x_date(date_labels = "%b %d") +
  scale_y_continuous(breaks=seq(0,10,by=0.5)) +
  coord_cartesian(ylim=c(0.75, 6)) +
  theme_bw() +
  theme(axis.text = element_text(colour="black"),
        axis.title = element_text(face="bold"),
        axis.title.x = element_blank(),
        legend.title = element_blank())
```

<br>

The last downloaded water level was **`r nitinat.water %>% filter(date2==max(date2)) %>% filter(hour==max(hour)) %>% pull(Water.Level.m)`m** on **`r nitinat.water %>% filter(date2==max(date2)) %>% filter(hour==max(hour)) %>% pull(date)`** at **`r nitinat.water %>% filter(date2==max(date2)) %>% filter(hour==max(hour)) %>% unite(col="time_stamp", time, stamp, sep=" ") %>% pull(time_stamp)`**. 

<br>

```{r fig.cap='Nitinat hydromet water and air temperature (C). Historical shaded area represents daily maximum and minimum recorded values over the time series, while the line represents the mean (gray). Current year line represents daily mean water and/or air temperature (green). Water and air temperature are recorded hourly at the Nitinat River Hatchery pumphouse.', fig.width=8, fig.height=6}

ggpubr::ggarrange(
  ggplot()+
    geom_ribbon(data=nitinat.water %>% 
                  filter(month %in% c("Aug","sep","Oct", "Nov") & year!=escapement_year) %>%
                  group_by(DOY) %>% 
                  summarize(max_temp=max(Water.temp.C,na.rm=T), min_temp=min(Water.temp.C,na.rm=T)),
                aes(x=as.Date(DOY, origin="2021-12-31"), ymin=min_temp, ymax=max_temp), fill="gray80") +
    geomtextpath:: geom_textline(data=nitinat.water %>% 
                                   filter(month %in% c("Aug","Sep","Oct", "Nov") & year!=escapement_year) %>%
                                   group_by(DOY) %>% 
                                   summarize(mean_temp_hist=mean(Water.temp.C,na.rm=T)) %>%
                                   mutate(label = paste0(min(nitinat.water$year), sep="-", escapement_year-1)),
                                 aes(x=as.Date(DOY, origin="2021-12-31"), y=mean_temp_hist, label=as.factor(label)), 
                                 colour="gray40", linewidth=1, show.legend=F, hjust=0.9, text_smoothing=30) +
    geomtextpath:: geom_labelline(data=nitinat.water %>% 
                                    filter(month %in% c("Aug","Sep","Oct", "Nov") & year==escapement_year) %>%
                                    group_by(DOY) %>% 
                                    summarize(mean_temp=mean(Water.temp.C,na.rm=T)) %>% 
                                    mutate(label=escapement_year),
                                  aes(x=as.Date(DOY, origin="2021-12-31"), y=mean_temp, label=as.factor(label)), 
                                  colour="green", linewidth=1, show.legend=F, hjust=0, text_smoothing=30) +
    labs(y="Water temperature (C)") +
    scale_colour_manual(values=c("green", "gray70")) +
    scale_x_date(date_labels = "%b %d") +
    #scale_y_continuous(breaks=seq(0,10,by=0.5)) +
    theme_bw() +
    theme(axis.text = element_text(colour="black"),
          axis.title = element_text(face="bold"),
          axis.title.x = element_blank(),
          legend.title = element_blank()),
  
  
  ggplot()+
    geom_ribbon(data=nitinat.water %>% 
                  filter(month %in% c("Aug","sep","Oct", "Nov") & year!=escapement_year) %>%
                  group_by(DOY) %>% 
                  summarize(max_temp=max(Air.Temp.Analog.C,na.rm=T), min_temp=min(Air.Temp.Analog.C,na.rm=T)),
                aes(x=as.Date(DOY, origin="2021-12-31"), ymin=min_temp, ymax=max_temp), fill="gray80") +
    geomtextpath:: geom_textline(data=nitinat.water %>% 
                                   filter(month %in% c("Aug","Sep","Oct", "Nov") & year!=escapement_year) %>%
                                   group_by(DOY) %>% 
                                   summarize(mean_temp_hist=mean(Air.Temp.Analog.C,na.rm=T)) %>%
                                   mutate(label = paste0(min(nitinat.water$year), sep="-", escapement_year-1)),
                                 aes(x=as.Date(DOY, origin="2021-12-31"), y=mean_temp_hist, label=as.factor(label)), 
                                 colour="gray40", linewidth=1, show.legend=F, hjust=0.91, text_smoothing=30) +
    geomtextpath:: geom_labelline(data=nitinat.water %>% 
                                    filter(month %in% c("Aug","Sep","Oct", "Nov") & year==escapement_year) %>%
                                    group_by(DOY) %>% 
                                    summarize(mean_temp=mean(Air.Temp.Analog.C,na.rm=T)) %>% 
                                    mutate(label=escapement_year),
                                  aes(x=as.Date(DOY, origin="2021-12-31"), y=mean_temp, label=as.factor(label)), 
                                  colour="green", linewidth=1, show.legend=F, hjust=0, text_smoothing=30) +
    labs(y="Air temperature (C)") +
    scale_colour_manual(values=c("green", "gray70")) +
    scale_x_date(date_labels = "%b %d") +
    #scale_y_continuous(breaks=seq(0,10,by=0.5)) +
    theme_bw() +
    theme(axis.text = element_text(colour="black"),
          axis.title = element_text(face="bold"),
          axis.title.x = element_blank(),
          legend.title = element_blank()),
  
  nrow=2)
```

<br>

<br>

## Environment Canada river forecast 

While there is no official flow forecast for the Nitinat River, the San Juan River Environment Canada COFFEE forecast can be used as a proxy for Nitinat River as both systems exhibit similar seasonal flow patterns.  

```{r}
knitr::include_url("http://bcrfc.env.gov.bc.ca/fallfloods/coffee/08HA010.PDF")
```

<br>

```{r fig.cap='Summer/fall water level at Nitinat (blue) and San Juan (gray) from 2021-present. Shaded region represents daily maximum and minimum levels while solid line indicates daily average among all years. San Juan water level from Environment Canada, Nitinat from South Coast Stock assessment hydromet.'}
ggplot()+
  geom_ribbon(data=SJ.water %>% 
                filter(month %in% c("Aug","Sep","Oct", "Nov") & year%in%c(2021:escapement_year) & PARAM=="level (m)")  %>%
                group_by(DOY) %>% 
                summarize(max_lvl=max(Value,na.rm=T), min_lvl=min(Value,na.rm=T)),
              aes(x=as.Date(DOY, origin="2021-12-31"), ymin=min_lvl, ymax=max_lvl), fill="gray80", alpha = 0.6) +
  geomtextpath::geom_textline(data=SJ.water %>% 
                                 filter(month %in% c("Aug","Sep","Oct", "Nov") & year%in%c(2021:escapement_year) & 
                                          PARAM=="level (m)")  %>%
                                 group_by(DOY) %>% 
                                 summarize(mean_lvl_hist=mean(Value,na.rm=T)) %>%
                                 mutate(label = paste0("San Juan ", min(nitinat.water$year), sep="-", escapement_year)),
                               aes(x=as.Date(DOY, origin="2021-12-31"), y=mean_lvl_hist, label=as.factor(label)), 
                               colour="gray60", linewidth=1, show.legend=F, hjust=0, text_smoothing=30, alpha=0.7) +
  geom_ribbon(data=nitinat.water %>% 
                filter(month %in% c("Aug","Sep","Oct", "Nov") & Water.Level.m>0.1) %>%
                group_by(DOY) %>% 
                summarize(max_lvl=max(Water.Level.m,na.rm=T), min_lvl=min(Water.Level.m,na.rm=T)),
              aes(x=as.Date(DOY, origin="2021-12-31"), ymin=min_lvl, ymax=max_lvl), fill="dodger blue", alpha=0.4) +
  geomtextpath:: geom_textline(data=nitinat.water %>% 
                                 filter(month %in% c("Aug","Sep","Oct", "Nov") & Water.Level.m>0.1) %>%
                                 group_by(DOY) %>% 
                                 summarize(mean_lvl=mean(Water.Level.m,na.rm=T)) %>% 
                                 mutate(label = paste0("Nitinat ", min(nitinat.water$year), sep="-", escapement_year)),
                               aes(x=as.Date(DOY, origin="2021-12-31"), y=mean_lvl, label=as.factor(label)), 
                               colour="dodger blue", linewidth=1, show.legend=F, hjust=0.01, alpha=0.8) +
  labs(y="Water level (m)") +
  scale_colour_manual(values=c("dodger blue", "gray70")) +
  scale_x_date(date_labels = "%b %d") +
  scale_y_continuous(breaks=seq(0,10,by=0.5)) +
  coord_cartesian(ylim=c(0.75, 6)) +
  theme_bw() +
  theme(axis.text = element_text(colour="black"),
        axis.title = element_text(face="bold"),
        axis.title.x = element_blank(),
        legend.title = element_blank())
```


<br>

<br>

# **Live and Dead Totals** 

<br> 

The `r escapement_year` Nitinat River and tribs surveys will begin on September 4. A preliminary swim was conducted prior to this to confirm Sockeye presence. See \@ref(tab:survey-table) for segments surveyed each date and \@ref(tab:peak-table) for peak counts. 

<br>

```{r survey-table, tab.cap='Segments surveyed by date.'}
nitinat.SIL %>% 
  filter(!is.na(date)) %>% 
  group_by(date, `upper/lower (for Nitinat River only)`) %>% 
  summarize() %>% 
  rename(`Segment(s) surveyed`=`upper/lower (for Nitinat River only)`,
         Date=date) %>% 
  kableExtra::kbl(align="c") %>%
  kableExtra::kable_paper("hover", full_width=T, position="center") 
```

<br>

<br>

Total counts:

```{r peak-table}
nitinat.SIL %>% 
  filter(!is.na(date), !grepl("no fish", species, ignore.case=T)) %>% 
  group_by(date, species, type) %>% 
  summarize(adults_live = sum(`adults live`)) %>% 
  kableExtra::kbl(align="c") %>%
  kableExtra::kable_paper("hover", full_width=T, position="center")
```

Peak counts:

```{r}
nitinat.SIL %>% 
  filter(!is.na(date), !grepl("no fish", species, ignore.case=T)) %>% 
  group_by(date, species, type) %>% 
  summarize(adults_live = sum(`adults live`)) %>% 
  group_by(species) %>% 
  filter(adults_live==max(adults_live)) %>%
  kableExtra::kbl(align="c") %>%
  kableExtra::kable_paper("hover", full_width=T, position="center")
```

<br>

<br>

```{r fig.cap='Raw live (bar) and carcass (line) counts for Nitinat River. Not all surveys are full system: early surveys are lower river only. Dead counts will populate once die-off begins.Note differing y-axis for each species.'}
ggplot() +
  # ---- DEAD -----
  geom_point(data=nitinat.SIL %>%
               filter(system=="Nitinat River" & species!="No fish") %>%
               group_by(date, species) %>%
               summarize(total_adults_ded=sum(`adults dead`)) ,
             aes(x=date, y=total_adults_ded, group=species, colour=species)) +
  # # ---- DEAD -----  
  geom_line(data=nitinat.SIL %>%
              filter(system=="Nitinat River" & species!="No fish") %>%
              group_by(date, species) %>%
              summarize(total_adults_ded=sum(`adults dead`)),
            aes(x=date, y=total_adults_ded, group=species, colour=species)) +
  # ---- LIVE ----
  geom_bar(data=nitinat.SIL %>% 
             filter(system=="Nitinat River" & species!="No fish") %>% 
             group_by(date, species, type) %>% 
             summarize(total_adults=sum(`adults live`)), 
           aes(x=date, y=total_adults, group=species, fill=species), stat="identity", alpha=0.8, width=2) +
  facet_wrap(~species, scales="free_y") +
  #scale_x_date(date_breaks="7 day", date_labels = "%b %d") +
  #scale_y_continuous(sec.axis = sec_axis(~ . /100, name = "Total dead adults (line)")) +
  labs(x="", y="Total live adults (bar)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        legend.position = "none", 
        axis.title = element_text(face="bold"))
```

<br>

<br>

```{r fig.cap='Live counts (bars) and water level (blue line) in the Nitinat River.'}
# Live ~ water 
# Plot counts with hydromet data to show pulses of fish into the system relative to rain events. 

ggplot() +
  # ---- WATER DATA ----
  geom_ribbon(data=nitinat.water %>% 
                filter(year==escapement_year & date2>=as.Date("2024-07-15")) %>%
                group_by(date2) %>% 
                summarize(mean_water = mean(Water.Level.m), min=min(Water.Level.m), max=max(Water.Level.m)),
              aes(x=date2, ymin=min*50, ymax=max*50), fill="dodger blue", alpha=0.5) +
  geom_line(data=nitinat.water %>% 
              filter(year==escapement_year & date2>=as.Date("2024-07-15")) %>%
              group_by(date2) %>% 
              summarize(mean_water = mean(Water.Level.m), min=min(Water.Level.m), max=max(Water.Level.m)),
            aes(x=date2, y=mean_water*50), colour="dodger blue", size=1) +
  # ---- COUNTS ----
  geom_bar(data=nitinat.SIL %>% 
             filter(system=="Nitinat River", !grepl("no fish", species, ignore.case=T)) %>% 
             group_by(date, species) %>% 
             summarize(total_adults=sum(`adults live / habitat seen (# Est Live)`)), 
           aes(x=date, y=total_adults, group=species, fill=species, colour=species), 
           stat="identity", alpha=0.8, width=2, size=1) +
  scale_x_date(date_breaks="7 day", date_labels = "%b %d") +
  scale_y_continuous(sec.axis = sec_axis(~ . /50, name = "water level (m)")) +
  labs(x="", y="Live count") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        legend.position = "none",
        axis.title = element_text(face="bold")) +
  facet_wrap(~species, ncol=2, scales="free_y")
```

<br>

<br>

# **Historical run timing** 

```{r fig.cap='Historical run timing for focal species in Nitinat River (gray). Stat weeks along x axis denote the month and week of historical counts (e.g., stat week 81=First week of August, statweek 93=Third week of Sept, etc). Current year counts will show as blue points when populated.', fig.width=12, fig.height=12}

ggplot() +
  geom_point(data=nit.rtStWk,
             aes(x=stat_week, y=count, colour=plot_group, fill=plot_group, size=plot_group),
             shape=21, alpha=0.6, stroke=1.5) +
  scale_fill_manual(breaks=waiver(), values=c("gray60", "dodger blue")) +
  scale_colour_manual(breaks=waiver(), values=c("gray60", "dodger blue")) +
  scale_size_manual(breaks=waiver(), values=c(2, 4)) +
  scale_y_continuous(labels=scales::comma) +
  labs(x="Stat week", y="Live adults (raw)", fill="Year:", colour="Year:", size="Year:"#, caption="Historical run timing for focal species in Nitinat River (gray). Stat weeks along x axis denote the month and week of historical counts (e.g., stat week 81=First week of August, \nstatweek 93=Third week of Sept, etc). Current year counts will show as blue points when populated. DFO South Coast Salmon Stock Assessment data."
       ) +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.box = "horizontal",
        legend.text = element_text(size=11),
        legend.title = element_text(face="bold", size=13),
        axis.title = element_text(face="bold", size=15),
        axis.text = element_text(size=12),
        legend.key.spacing.x = unit(5,"mm"),
        plot.caption = element_text(hjust=0)) +
  facet_wrap(~stringr::str_to_title(species), nrow=3, scales="free_y") 
```

```{r pdfrt, eval=F}
pdf(file = here::here("outputs", "figures", "Nitinat run timing.pdf"),   
    width = 11, # The width of the plot in inches
    height = 8.5) # The height of the plot in inches

ggplot() +
  #annotate("text", x=Inf, y=-Inf, label="NOT FINAL ESTIMATES. DO NOT SHARE.",
  #         hjust=1.1, vjust=-1.1, col="gray70", size=6, fontface="bold", alpha=0.6) +
  geom_point(data=nit.rtStWk,
             aes(x=stat_week, y=count, colour=plot_group, fill=plot_group, size=plot_group),
             shape=21, alpha=0.6, stroke=1.5) +
  scale_fill_manual(breaks=waiver(), values=c("gray60", "dodger blue")) +
  scale_colour_manual(breaks=waiver(), values=c("gray60", "dodger blue")) +
  scale_size_manual(breaks=waiver(), values=c(2, 4)) +
  scale_y_continuous(labels=scales::comma) +
  labs(x="Stat week", y="Live adults (raw)", fill="Year:", colour="Year:", size="Year:", caption="Historical run timing for focal species in Nitinat River (gray). Stat weeks along x axis denote the month and week of historical counts (e.g., stat week 81=First week of August, \nstatweek 93=Third week of Sept, etc). Current year counts will show as blue points when populated. DFO South Coast Salmon Stock Assessment data.") +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.box = "horizontal",
        axis.title = element_text(face="bold"),
        legend.key.spacing.x = unit(5,"mm"),
        plot.caption = element_text(hjust=0)) +
  facet_wrap(~stringr::str_to_title(species), nrow=3, scales="free_y") 

dev.off()
```


<br>

<br>

```{r fig.cap='Historical run timing for focal species in Nitinat River (gray). Stat weeks along x axis denote the month and week of historical counts (e.g., stat week 81=First week of August, statweek 93=Third week of Sept, etc). Current year counts will show as blue points when populated.', fig.width=12, fig.height=12}
#
cowplot::plot_grid(

ggplot() +
  geom_tile(data=nit.rtStWk %>%
              filter(species=="chinook") %>%
              group_by(year, stat_week) %>%
              summarize(count=case_when(!is.na(count) ~ sum(count,na.rm=T),
                                        is.na(count) ~ NA), plot_group=plot_group) %>%
              filter(!is.na(count)),
            aes(x=stat_week, y=reorder(year, year, descending=F), fill=count)) +
  scale_fill_viridis_c(option = "C") +
  scale_y_discrete(limits=rev) +
  labs(x="Stat week", y="", fill="Chinook\nLive adults: ") +
  theme_bw() +
  theme(legend.position = c(0.93, 0.2),
        legend.text = element_text(size=9),
        legend.title = element_text(size=11),
        legend.background = element_rect(fill=alpha("transparent",0.7), colour="black"),
        axis.title = element_text(face="bold", size=15),
        axis.text = element_text(colour="black", size=12),
        legend.key.spacing.x = unit(5,"mm"),
        plot.caption = element_text(hjust=0)),

ggplot() +
  geom_tile(data=nit.rtStWk %>%
              filter(species=="coho") %>%
              group_by(year, stat_week) %>%
              summarize(count=case_when(!is.na(count) ~ sum(count,na.rm=T),
                                        is.na(count) ~ NA), plot_group=plot_group) %>%
              filter(!is.na(count)),
            aes(x=stat_week, y=reorder(year, year, descending=F), fill=count)) +
  scale_fill_viridis_c(option = "C") +
  scale_y_discrete(limits=rev) +
  labs(x="Stat week", y="", fill="Coho\nLive adults: ") +
  theme_bw() +
  theme(legend.position = c(0.93, 0.22),
        legend.text = element_text(size=9),
        legend.title = element_text(size=11),
        legend.background = element_rect(fill=alpha("transparent",0.7), colour="black"),
        axis.title = element_text(face="bold", size=15),
        axis.text = element_text(colour="black", size=12),
        legend.key.spacing.x = unit(5,"mm"),
        plot.caption = element_text(hjust=0)),

ggplot() +
  geom_tile(data=nit.rtStWk %>%
              filter(species=="chum") %>%
              group_by(year, stat_week) %>%
              summarize(count=case_when(!is.na(count) ~ sum(count,na.rm=T),
                                        is.na(count) ~ NA), plot_group=plot_group) %>%
              filter(!is.na(count)),
            aes(x=stat_week, y=reorder(year, year, descending=F), fill=count)) +
  scale_fill_viridis_c(option = "C", guide = guide_colorbar(barwidth = 0.5, barheight = 5)) +
  scale_y_discrete(limits=rev) +
  labs(x="Stat week", y="", fill="Chum\nLive adults: ") +
  theme_bw() +
  theme(legend.position = c(0.93, 0.45),
        legend.text = element_text(size=9),
        legend.title = element_text(size=11),
        legend.background = element_rect(fill=alpha("transparent",0.7), colour="black"),
        axis.title = element_text(face="bold", size=15),
        axis.text = element_text(colour="black", size=12),
        legend.key.spacing.x = unit(5, "mm"),
        plot.caption = element_text(hjust=0)),

nrow=2)
```

```{r pdfheatmap, eval=F}
# PDF VERSION 

pdf(file = here::here("outputs", "figures", "Nitinat run timing (heatmap).pdf"),   
    width = 11, # The width of the plot in inches
    height = 8.5) # The height of the plot in inches

rtheat <- cowplot::plot_grid(

ggplot() +
  geom_tile(data=nit.rtStWk %>%
              filter(species=="chinook") %>%
              group_by(year, stat_week) %>%
              summarize(count=case_when(!is.na(count) ~ sum(count,na.rm=T),
                                        is.na(count) ~ NA), plot_group=plot_group) %>%
              filter(!is.na(count)),
            aes(x=stat_week, y=reorder(year, year, descending=F), fill=count)) +
  scale_fill_viridis_c(option = "C") +
  scale_y_discrete(limits=rev) +
  labs(x="Stat week", y="", fill="Chinook\nLive adults: ") +
  theme_bw() +
  theme(legend.position = c(0.93, 0.22),
        legend.text = element_text(size=6),
        legend.title = element_text(size=7),
        legend.background = element_rect(fill=alpha("transparent",0.7), colour="black"),
        axis.title = element_text(face="bold"),
        axis.text = element_text(colour="black"),
        legend.key.spacing.x = unit(5,"mm"),
        plot.caption = element_text(hjust=0)),

ggplot() +
  geom_tile(data=nit.rtStWk %>%
              filter(species=="coho") %>%
              group_by(year, stat_week) %>%
              summarize(count=case_when(!is.na(count) ~ sum(count,na.rm=T),
                                        is.na(count) ~ NA), plot_group=plot_group) %>%
              filter(!is.na(count)),
            aes(x=stat_week, y=reorder(year, year, descending=F), fill=count)) +
  scale_fill_viridis_c(option = "C") +
  scale_y_discrete(limits=rev) +
  labs(x="Stat week", y="", fill="Coho\nLive adults: ") +
  theme_bw() +
  theme(legend.position = c(0.93, 0.22),
        legend.text = element_text(size=6),
        legend.title = element_text(size=7),
        legend.background = element_rect(fill=alpha("transparent",0.7), colour="black"),
        axis.title = element_text(face="bold"),
        axis.text = element_text(colour="black"),
        legend.key.spacing.x = unit(5,"mm"),
        plot.caption = element_text(hjust=0)),

ggplot() +
  geom_tile(data=nit.rtStWk %>%
              filter(species=="chum") %>%
              group_by(year, stat_week) %>%
              summarize(count=case_when(!is.na(count) ~ sum(count,na.rm=T),
                                        is.na(count) ~ NA), plot_group=plot_group) %>%
              filter(!is.na(count)),
            aes(x=stat_week, y=reorder(year, year, descending=F), fill=count)) +
  scale_fill_viridis_c(option = "C", guide = guide_colorbar(barwidth = 0.5, barheight = 5)) +
  scale_y_discrete(limits=rev) +
  labs(x="Stat week", y="", fill="Chum\nLive adults: ") +
  theme_bw() +
  theme(legend.position = c(0.93, 0.22),
        legend.text = element_text(size=6),
        legend.title = element_text(size=7),
        legend.background = element_rect(fill=alpha("transparent",0.7), colour="black"),
        axis.title = element_text(face="bold"),
        axis.text = element_text(colour="black"),
        legend.key.spacing.x = unit(5,"mm"),
        plot.caption = element_text(hjust=0)),

nrow=2)

dev.off()
```













