---
title: "Area 21/121 Fishery Summary"
date: 'Last update: `r format(Sys.time(), sep="_", "%Y-%m-%d %H:%M")`'
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_collapsed: false 
    toc_depth: 5
    number_sections: true
---

```{r setup, include=F}
# title: "part3-fisheries-knitable"

knitr::opts_chunk$set(echo=F,  message=F, warning=F)


# Load libraries ------------------
library(tidyverse)
# library(here)
# library(readxl)
# library(scales)
# library(plotly)
# library(bookdown)
# library(kableExtra)


# Load helpers ------------------
"%notin%" <- Negate("%in%")








# ============================ Load data ============================

# Biodata w results ----------------------
WCVIcrestBDWR <- readxl::read_excel(path=list.files(path=#"//ENT.dfo-mpo.ca/DFO-MPO/GROUP/PAC/PBS/Operations/SCA/SCD_Stad/WCVI/CHINOOK/WCVI_TERMINAL_RUN/Annual_data_summaries_for_RunRecons/CREST-BDWRcompile_base-files/2-Export-from-R/",
                                            here::here("data", "fisheries"),
                                          pattern="^R_OUT - Biological_Data_With_Results \\(.*\\) \\d{4}-\\d{4}\\.xlsx$", 
                                          full.names=T), 
                          sheet="CREST Biodata Compiled", guess_max = 10000) %>%
  mutate(yday = lubridate::yday(COLLECTION_DATE),
         LEGAL_STATUS = case_when(LENGTH_MM > 450 & LENGTH_MM < 800 ~ "LEGAL",
                                  LENGTH_MM <= 450 ~ "SUBLEGAL",
                                  LENGTH_MM >= 800 & yday%in%c(196:212) ~ "SUPERLEGAL",
                                  LENGTH_MM >= 800 & yday%notin%c(196:212) ~ "LEGAL"),
         isNit1 = case_when(grepl("Nitinat", RESOLVED_STOCK_ORIGIN) ~ paste0(`(R) Origin`, " ", RESOLVED_STOCK_ORIGIN),
                           !grepl("Nitinat", RESOLVED_STOCK_ORIGIN) & `(R) Origin`!="Unknown" ~ 
                             paste0(`(R) Origin`, " Other"),
                           !grepl("Nitinat", RESOLVED_STOCK_ORIGIN) & `(R) Origin`=="Unknown" ~ 
                             "Unknown Other"),
         isNit2 = case_when(grepl("Nitinat", isNit1) ~ "Nitinat (all)",
                            TRUE ~ "Other"),
         isNit3 = case_when(isNit1=="Hatchery Nitinat River" ~ "Nitinat (hatchery)",
                             TRUE ~ "Other")) %>% 
  filter(!(RESOLVED_STOCK_SOURCE=="DNA" & PROB_1<0.75)) %>%
  print()


# Catch estimates ----------------------
catEst <- readxl::read_excel(path="//ENT.dfo-mpo.ca/DFO-MPO/GROUP/PAC/Reg_Shares/FHM/SMCS/Salmon/FMCR_Fishery_Monitoring_Catch_Reporting/Recreational_CM/Catch_Data/SC Sport Catch Creel Sub-area Disposition (Master Do No Edit).xlsx", 
                             sheet="YTD") %>%
  filter(grepl("127|27|126|26|125|25|124|24|123|23|121|21|22|20", PFMA), YEAR>=2017, SPECIES=="CHINOOK SALMON") 




# CREST biodata joined to catch ------------------------
WCVIcrestBDWR.catEst.Nit <- left_join(WCVIcrestBDWR %>% 
                                        mutate_at("AREA", as.character) %>% 
                                        filter(SAMPLE_TYPE=="Sport", SPECIES==124, DISPOSITION=="Kept", LEGAL_STATUS=="LEGAL") %>% 
                                        group_by(YEAR, MONTH, AREA, isNit1, isNit2, isNit3) %>% 
                                        summarize(n_lvl1 = n()), 
                                      
                                      catEst %>% 
                                        mutate(PFMA = str_sub(PFMA, start=6, end=9)) %>%
                                        group_by(YEAR, PFMA, MONTH, DISPOSITION) %>% 
                                        summarize(Est = sum(ESTIMATE, na.rm=T),
                                                  SE = sum(STANDARD_ERROR, na.rm=T)) %>% 
                                        pivot_wider(names_from=DISPOSITION, values_from=c(Est, SE)) %>%
                                        select(YEAR, PFMA, MONTH, Est_Kept, `Est_Released Legal`),
                                      
                                      by=c("YEAR", 
                                           "MONTH", 
                                           "AREA" = "PFMA")) %>%
  full_join(data.frame(AREA=c("127", "27", "126", "26", "125", "25", "124", "24", "123", "23", "121", "22", "21", "20")),
            .,
            by="AREA") %>%
  print()




# ================================= Spatial data =================================

# DFO PFMA spatial files -----------------------
DFOcentroids <- read.csv(here::here("data", "fisheries", "spatial", "MRP_PFMA_spatial.csv")) %>%
  rename(`R (RC) Recovery PSC Location-L4`=PFMA_resolved)

pfmaPOLY <- rgdal::readOGR(dsn=here::here("data", "fisheries", "spatial"), layer="pfma1", verbose=F)
wcviPFMApoly <- pfmaPOLY[pfmaPOLY@data$STATAREA%in%c(20:27,121:127),]
pfma_sf <- sf::st_read(dsn=here::here("data", "fisheries", "spatial"), layer="pfma1")

creelSubAreaPOLY <- rgdal::readOGR(dsn=here::here("data", "fisheries", "spatial"), layer="Creel_Survey_Areas", verbose=F)
creelSubAreaPOLY_sf <- sf::st_read(dsn=here::here("data", "fisheries", "spatial"), layer="Creel_Survey_Areas")
creelSubAreaPOLY_sf <- sf::st_as_sf(x = creelSubAreaPOLY_sf,                         
                  coords = c("longitude", "latitude"),
                  crs = "+proj=utm +zone=10")
sfc = sf::st_transform(creelSubAreaPOLY_sf, crs = "+proj=longlat +datum=WGS84")
creelSubAreaPOLY_sfdf <- as.data.frame(sfc)
  
# Load "UScentroids" -----------------------
#source(here("scripts", "misc-helpers", "UScentroidCompile.R"))




# Join Catch Est data to Spatial data ------------------
wcviPFMA_NitCatEst <- left_join(pfma_sf,
                                
                               WCVIcrestBDWR.catEst.Nit %>% 
                                 group_by(YEAR, MONTH, AREA, isNit3) %>% 
                                 summarize(n_lvl3 = sum(n_lvl1, na.rm=T),
                                           Est_Kept = unique(Est_Kept),
                                           `Est_Released Legal` = unique(`Est_Released Legal`))  %>%
                                 group_by(YEAR, MONTH, AREA) %>% 
                                 mutate(AreaMonthTotalSamples = sum(n_lvl3, na.rm=T),
                                        propn_AreaMonth = n_lvl3/AreaMonthTotalSamples,
                                        estKeptByID = propn_AreaMonth*Est_Kept,
                                        NitOnly_forPlot = case_when(grepl("Other", isNit3) ~ Est_Kept-estKeptByID,
                                                                    TRUE ~ estKeptByID),
                                        NitOnly_forPlot = case_when(NitOnly_forPlot==0 ~ NA,
                                                                    TRUE ~ NitOnly_forPlot)) %>%
                                 filter(!is.na(YEAR),
                                        !(isNit3=="Other" & propn_AreaMonth<1 & estKeptByID!=NitOnly_forPlot)) %>%
                                 ungroup() %>%
                                 full_join(data.frame(AREA=c("127", "27", "126", "26", "125", "25", "124", "24", 
                                                             "123", "23", "121", "22", "21", "20")) %>%
                                             mutate(YEAR=rep(2023, nrow(.)),
                                                    MONTH=rep("May", nrow(.))),
                                           .,
                                           by=c("YEAR", "MONTH", "AREA")) %>%
                                 complete(., YEAR, MONTH, AREA) %>% 
                                 ungroup() %>%
                                 mutate_at("AREA", as.character) %>%
                                 mutate_at("AREA", as.integer), 
                               
                               by=c("STATAREA"="AREA")
)

```

<br> 



# **Sport Fishery Biodata Results**

<br>

## **Assumptions and caveats**

Definitions of natural, hatchery and unknown origin Chinook:

- A fish is assumed Natural origin if an otolith is returned as "Not Marked" and the adipose fin is intact. PBT is not currently considered as it is stock/facility-specific, and would likely only apply to very recent RYs (2022/2023). 
  + This assumption may not be appropriate in cases where hatchery fish are not thermally marked AND have no other identifier (probably rare? Perhaps US?).
  + Intact adipose fins *alone* are NOT evidence of a natural fish, and are considered "Unknown" origin unless accompanied by more data
- A fish is Hatchery origin if it has an adipose clip, a thermal mark, a CWT, and/or a PBT BY assignment (in recent years).
  + Note there are some rare cases of CWT wild-tagging programs in Yukon/Alaska, but these are not expected to show up in great numbers for Chinook. 
- A fish is Unknown origin if it has an intact adipose and no other mark or tag to inform further, e.g., destroyed or missing otolith, missing or lost tag, no  samples analyzed, etc.) 
- Stock ID is assigned in order of reliability: CWT > PBT > Otolith > GSI. This is determined in the CREST database but is currently under review (PBT may shift to be more reliable). 

<br>

The following caveats/assumptions must be taken into consideration when interpreting the recreational fishery data presented below:

- A genetic stock ID is only included if it has >= 75% certainty.
- It is assumed that biosampling is representative of kept catch, but this varies year to year as sample sizes are low. 
- Biosamples represent kept, legal catch and therefore do not represent the portions of populations <45cm or >80cm
  + Only legal sized Chinook are included here (> 45cm). Extra-legal (aka Super-legal) Chinook (>80cm) have also been excluded as current regs exclude these fish from the fishery.
- **No corrections are done to account for thermal mark issues from any facility in any given year. Missing  thermal marks would show up as "Not Marked" and bias results towards reporting more natural-origin Chinook than may in fact exist.**
- Limimted GSI sampling in earlier years
- Considerable changes to the recreational fishery along SWVI came into effect around 2018-2019 so comparisons across these years should be careful. 
- Unknown stock IDs were excluded from the following analyses. Results are expressed as regional population IDs as a proportion of known samples. 

<br>

**Please keep in mind sample rates for these areas are below ideal levels, so conclusions drawn should be very careful. Please read the last section that lays out the sample rates (i.e., % of kept catch sampled) by month and year!"**

<br>

<br>

## **WCVI: Nitinat catch across all WCVI PFMAs (27-20 and offshore)** 

Proportion of legal sized biosamples that return as Nitinat in each Area applied to kept, legal catch estimates for 2017-2023 (note 2023 biodata still incomplete). 

<br>

### **All Nitinat-origin across WCVI (hatchery + natural)**

Applying stock comp %s from creel biosamples to estimated catch by Month/Area/Year. Note 2023 and 2024 are incomplete as of January 2025. 

```{r nitHeat, out.width='1200px', out.height='700px', fig.cap='Kept catch estimates of legal size (>45cm and <=80cm) Nitinat-origin Chinook across WCVI PFMAs, organized from north (PFMA 127) to south (PFMA 20). Estimates calculated by applying the proportion of Nitinat chinook detected in creel biosamples to total year-month-area catch estimates. Interactive plot; hover over data points on heat map to see sample sizes and catch estimates. **NOTE 2023 and 2024 samples are incomplete as of January 2025.**'}

# Heatmap of all WCVI Nitinat catch estimates all years ------------------


# pdf(file = here::here("outputs", "figures", "WCVI Nitinat catch estimate heatmap 2017-2023.pdf"),   
#     width = 11, # The width of the plot in inches
#     height = 8.5) # The height of the plot in inches

plotly::ggplotly(
  ggplot(data = WCVIcrestBDWR.catEst.Nit %>% 
           group_by(YEAR, MONTH, AREA, isNit2) %>% 
           summarize(n_lvl2 = sum(n_lvl1, na.rm=T),
                     Est_Kept = unique(Est_Kept),
                     `Est_Released Legal` = unique(`Est_Released Legal`))  %>%
           group_by(YEAR, MONTH, AREA) %>% 
           mutate(AreaMonthTotalSamples = sum(n_lvl2, na.rm=T),
                  propn_AreaMonth = n_lvl2/AreaMonthTotalSamples,
                  estKeptByID = propn_AreaMonth*Est_Kept,
                  NitOnly_forPlot = case_when(grepl("Other", isNit2) ~ Est_Kept-estKeptByID,
                                              TRUE ~ estKeptByID),
                  NitOnly_forPlot = case_when(NitOnly_forPlot==0 ~ NA,
                                              TRUE ~ NitOnly_forPlot)) %>%
           filter(!is.na(YEAR),
                  !(grepl("Other", isNit2) & propn_AreaMonth<1 & estKeptByID!=NitOnly_forPlot)) %>%
           ungroup() %>%
           full_join(data.frame(AREA=c("127", "27", "126", "26", "125", "25", "124", "24", "123", "23", "121", "22", "21", "20")) %>%
                       mutate(YEAR=rep(2023, nrow(.)),
                              MONTH=rep("May", nrow(.))),
                     .,
                     by=c("YEAR", "MONTH", "AREA")) %>%
           complete(., YEAR, MONTH, AREA)) +
    # geom_rect(data=WCVIcrestBDWR.catEst.Nit %>% 
    #           filter(YEAR %in% c(2023, 2024)) %>%
    #           group_by(YEAR) %>%
    #           summarize(unique(YEAR)), 
    #         aes(), fill="red", xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf, alpha=0.1) +
    geom_tile(aes(y=AREA, x=factor(MONTH, levels=month.name), fill=NitOnly_forPlot, colour=NitOnly_forPlot,
                  text=paste("Estimated # kept Nitinat CN=", round(estKeptByID,0), "\nBased on # confirmed biosamples, n=", n_lvl2)), 
              stat="identity", alpha=0.9) +
    paletteer::scale_fill_paletteer_c("viridis::plasma", na.value = alpha("white", 0.1)) +
    paletteer::scale_colour_paletteer_c("viridis::plasma", na.value = alpha("gray80", 0.3)) +
    scale_x_discrete(limits = c("May", "June", "July", "August", "September"), 
                     labels=c("May", "Jun", "Jul", "Aug", "Sept")) +
    scale_y_discrete(limits = rev(c("127", "27", "126", "26", "125", "25", "124", "24", "123", "23", "121", "22", "21", "20"))) +
    labs(x="", y="Area", fill="Estimated # \nkept legal \nNitinat-origin CN", 
         colour="Estimated # \nkept legal \nNitinat-origin CN") +
    theme_bw() +
    facet_wrap(~YEAR, nrow=1) + 
    theme(axis.title = element_text(face="bold", size=10),
          axis.text.y = element_text(colour="black", size=8),
          axis.text.x = element_text(colour="black", size=9, angle=45, hjust=1),
          legend.title = element_text(face="bold", size=12),
          legend.text = element_text(size=8),
          #legend.key.size = unit(5, "mm"),
          #legend.position = "bottom",
          #legend.direction = "horizontal",
          strip.text = element_text(size=10)),
  
  tooltip = c("AREA", "MONTH", "text")
)

#dev.off()
```


<br>

<br>


### **Hatchery Nitinat across WCVI** 

<br>

#### 2023 Catch estimate map

```{r wcvi23, out.width='1000px', out.height='600px', fig.cap='Kept catch estimates of legal sized (>45cm and <=80cm) Nitinat Hatchery Chinook across WCVI in 2023. Estimates calculated by applying the proportion of Nitinat chinook detected in creel biosamples to total year-month-area catch estimates. **NOTE 2023 and 2024 samples are incomplete as of January 2025.**'}

#pdf(file = here::here("outputs", "figures", "WCVI Nitinat catch estimate map 2023.pdf"),   
#    width = 11, # The width of the plot in inches
#    height = 8.5) # The height of the plot in inches

#htmltools::save_html(
#plotly::ggplotly(
ggplot() +
  geom_sf(data=wcviPFMA_NitCatEst %>% 
            sf::st_set_crs(.,"WGS84") %>% 
            filter(YEAR=="2023") %>%
            filter(MONTH%in%c("May", "June", "July", "August", "September")), 
          aes(fill=round(NitOnly_forPlot, 0), 
              text=paste("Estimated # kept Nitinat CN=", round(NitOnly_forPlot,0), 
                         "\nBased on # confirmed biosamples, n=", n_lvl3)), 
          colour=alpha("gray60", 0.5), linewidth=0.5) +
  paletteer::scale_fill_paletteer_c("viridis::plasma", na.value = NA) +
  geom_sf(data=rnaturalearth::ne_countries(type="countries", scale="large", returnclass="sf"), 
          colour="transparent", fill="black") +
  coord_sf(crs="+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0", xlim=c(-129, -123), ylim=c(51.5, 48)) +
  labs(fill="Estimated number of kept legal \nHatchery Nitinat Chinook 2023") +
  theme_bw() +
  theme(axis.text = element_blank(),
        legend.title = element_text(face="bold", size=10),
        legend.text = element_text(size=9),
        legend.position = c(0.85, 0.2),
        panel.grid.major = element_blank(),
        panel.background = element_rect(fill="white")) +
  facet_wrap(~factor(MONTH, levels=c("May", "June", "July", "August", "September", ordered=T)))
#,
#  tooltip = c("text")
#),

#file = here::here("outputs", "figures", "WCVI Nitinat catch estimate map 2023.html")
#)


#dev.off()
```

<br>

#### 2022 Catch estimate map

```{r wcvi22, out.width='1000px', out.height='600px', fig.cap='Kept catch estimates of legal sized (>45cm and <=80cm) Nitinat Hatchery Chinook across WCVI in 2022. Estimates calculated by applying the proportion of Nitinat chinook detected in creel biosamples to total year-month-area catch estimates.'}
#pdf(file = here::here("outputs", "figures", "WCVI Nitinat catch estimate map 2022.pdf"),   
#    width = 11, # The width of the plot in inches
#    height = 8.5) # The height of the plot in inches

#plotly::ggplotly(
ggplot() +
  geom_sf(data=wcviPFMA_NitCatEst %>% 
            sf::st_set_crs(.,"WGS84") %>% 
            filter(YEAR=="2022") %>%
            filter(MONTH%in%c("May", "June", "July", "August", "September")), 
          aes(fill=round(NitOnly_forPlot, 0), 
              text=paste("Estimated # kept Nitinat CN=", round(NitOnly_forPlot,0), 
                         "\nBased on # confirmed biosamples, n=", n_lvl3)), 
          colour=alpha("gray60", 0.5), linewidth=0.5) +
  paletteer::scale_fill_paletteer_c("viridis::plasma", na.value = NA) +
  geom_sf(data=rnaturalearth::ne_countries(type="countries", scale="large", returnclass="sf"), 
          colour="transparent", fill="black") +
  coord_sf(crs="+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0", xlim=c(-129, -123), ylim=c(51.5, 48)) +
  labs(fill="Estimated number of kept legal \nHatchery Nitinat Chinook 2022") +
  theme_bw() +
  theme(axis.text = element_blank(),
        legend.title = element_text(face="bold", size=10),
        legend.text = element_text(size=9),
        legend.position = c(0.85, 0.2),
        panel.grid.major = element_blank(),
        panel.background = element_rect(fill="white")) +
  facet_wrap(~factor(MONTH, levels=c("May", "June", "July", "August", "September", ordered=T)))
#,
#  tooltip = c("text")
#)

#dev.off()
```

<br>

#### 2021 Catch estimate map

```{r wcvi21, out.width='1000px', out.height='600px', fig.cap='Kept catch estimates of legal sized (>45cm and <=80cm) Nitinat Hatchery Chinook across WCVI in 2021. Estimates calculated by applying the proportion of Nitinat chinook detected in creel biosamples to total year-month-area catch estimates.'}


#pdf(file = here::here("outputs", "figures", "WCVI Nitinat catch estimate map 2021.pdf"),   
#    width = 11, # The width of the plot in inches
#    height = 8.5) # The height of the plot in inches

#plotly::ggplotly(
ggplot() +
  geom_sf(data=wcviPFMA_NitCatEst %>% 
            sf::st_set_crs(.,"WGS84") %>% 
            filter(YEAR=="2021") %>%
            filter(MONTH%in%c("May", "June", "July", "August", "September")), 
          aes(fill=round(NitOnly_forPlot, 0), 
              text=paste("Estimated # kept Nitinat CN=", round(NitOnly_forPlot,0), 
                         "\nBased on # confirmed biosamples, n=", n_lvl3)), 
          colour=alpha("gray60", 0.5), linewidth=0.5) +
  paletteer::scale_fill_paletteer_c("viridis::plasma", na.value = NA) +
  geom_sf(data=rnaturalearth::ne_countries(type="countries", scale="large", returnclass="sf"), 
          colour="transparent", fill="black") +
  coord_sf(crs="+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0", xlim=c(-129, -123), ylim=c(51.5, 48)) +
  labs(fill="Estimated number of kept legal \nHatchery Nitinat Chinook 2021") +
  theme_bw() +
  theme(axis.text = element_blank(),
        legend.title = element_text(face="bold", size=10),
        legend.text = element_text(size=9),
        legend.position = c(0.85, 0.2),
        panel.grid.major = element_blank(),
        panel.background = element_rect(fill="white")) +
  facet_wrap(~factor(MONTH, levels=c("May", "June", "July", "August", "September", ordered=T)))
#,
#  tooltip = c("text")
#)

#dev.off()
```

<br>

#### 2020 Catch estimate map

```{r wcvi20, out.width='1000px', out.height='600px', fig.cap='Kept catch estimates of legal sized (>45cm and <=80cm) Nitinat Hatchery Chinook across WCVI in 2020. Estimates calculated by applying the proportion of Nitinat chinook detected in creel biosamples to total year-month-area catch estimates.'}


#pdf(file = here::here("outputs", "figures", "WCVI Nitinat catch estimate map 2020.pdf"),   
#    width = 11, # The width of the plot in inches
#    height = 8.5) # The height of the plot in inches

#plotly::ggplotly(
ggplot() +
  geom_sf(data=wcviPFMA_NitCatEst %>% 
            sf::st_set_crs(.,"WGS84") %>% 
            filter(YEAR=="2020") %>%
            filter(MONTH%in%c("May", "June", "July", "August", "September")), 
          aes(fill=round(NitOnly_forPlot, 0), 
              text=paste("Estimated # kept Nitinat CN=", round(NitOnly_forPlot,0), 
                         "\nBased on # confirmed biosamples, n=", n_lvl3)), 
          colour=alpha("gray60", 0.5), linewidth=0.5) +
  paletteer::scale_fill_paletteer_c("viridis::plasma", na.value = NA) +
  geom_sf(data=rnaturalearth::ne_countries(type="countries", scale="large", returnclass="sf"), 
          colour="transparent", fill="black") +
  coord_sf(crs="+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0", xlim=c(-129, -123), ylim=c(51.5, 48)) +
  labs(fill="Estimated number of kept legal \nHatchery Nitinat Chinook 2020") +
  theme_bw() +
  theme(axis.text = element_blank(),
        legend.title = element_text(face="bold", size=10),
        legend.text = element_text(size=9),
        legend.position = c(0.85, 0.2),
        panel.grid.major = element_blank(),
        panel.background = element_rect(fill="white")) +
  facet_wrap(~factor(MONTH, levels=c("May", "June", "July", "August", "September", ordered=T)))
#,
#  tooltip = c("text")
#)

#dev.off()
```

<br>

<br>


















## **Area 21/121 catch composition**

```{r echo=F, message=F, warning=F}

# Join ALL BDWR  to catch and calculate # caught by ID (for Area 21/121) -----------------
WCVIcrestBDWR.catEst <- left_join(WCVIcrestBDWR %>%
                                    filter(SAMPLE_TYPE=="Sport", SPECIES==124, DISPOSITION=="Kept", LEGAL_STATUS=="LEGAL") %>% 
                                    group_by(YEAR, MONTH, AREA, RESOLVED_STOCK_ORIGIN, RESOLVED_STOCK_ROLLUP) %>% 
                                    summarize(n_origin = n()),
                                  catEst %>% 
                                    filter(SPECIES_CODE=="124") %>%
                                    mutate(PFMA = str_sub(PFMA, start=6, end=9)) %>%
                                    group_by(YEAR, PFMA, MONTH, DISPOSITION) %>% 
                                    summarize(Est = sum(ESTIMATE),
                                              `%SE` = sum(PERCENT_STANDARD_ERROR)) %>% 
                                    pivot_wider(names_from=DISPOSITION, values_from=c(Est, `%SE`)) %>%
                                    select(YEAR, PFMA, MONTH, Est_Kept, `%SE_Kept`, `Est_Released Legal`, `%SE_Released Legal`, 
                                           `Est_Released Sub-legal`, `%SE_Released Sub-legal`),
                                  by=c("AREA"="PFMA", "YEAR", "MONTH"))  %>%
  mutate(NitRegionID = case_when(grepl("Nitinat", RESOLVED_STOCK_ORIGIN) ~ RESOLVED_STOCK_ORIGIN,
                                 TRUE ~ RESOLVED_STOCK_ROLLUP)) %>%
  relocate(NitRegionID, .after=RESOLVED_STOCK_ROLLUP) 
```


```{r a21comp, out.width='1300px', out.height='500px', fig.align='left', fig.cap='Estimated kept, legal (<45cm, >80cm) Chinook catch in Area 21. ID is to the Region-level in all cases except for Nitinat River. Estimates calculated by applying the proportion of Nitinat chinook detected in creel biosamples to total year-month-area catch estimates. **NOTE 2023 and 2024 samples in incomplete as of January 2025.**'}

# Region rollup stock comp --------------------

# AREA 21
plotly::ggplotly(
  ggplot() +
    geom_tile(data=WCVIcrestBDWR.catEst %>% 
                filter(AREA=="21") %>% 
                group_by(YEAR, MONTH, AREA, NitRegionID) %>% 
                summarize(nID = sum(n_origin, na.rm=T),
                          TotalCatch = unique(Est_Kept)) %>% 
                group_by(YEAR, MONTH, AREA) %>% 
                mutate(totalSamples = sum(nID),
                       propn_NitRegionID = nID/totalSamples,
                       catchByRegionID = round(propn_NitRegionID*TotalCatch,0)) %>% 
                ungroup() %>%
                filter(!is.na(NitRegionID)) %>%
                full_join(data.frame(expand.grid(NitRegionID=WCVIcrestBDWR.catEst %>% 
                                                   filter(AREA=="21") %>% 
                                                   group_by(NitRegionID) %>% 
                                                   summarize(NitRegionID=unique(NitRegionID)) %>% 
                                                   filter(!is.na(NitRegionID)) %>%
                                                   pull(NitRegionID),
                                                 YEAR = c(min(WCVIcrestBDWR.catEst$YEAR):max(WCVIcrestBDWR.catEst$YEAR)),
                                                 MONTH = c("May","June","July","August","September"),
                                                 AREA = "21")),
                          .,
                          by=c("YEAR", "MONTH", "AREA", "NitRegionID")) %>%
                complete(., YEAR, MONTH, AREA, NitRegionID) %>%
                filter(!is.na(NitRegionID)) %>%
                rename(`Regional ID`=NitRegionID)
              ,
              aes(x=factor(MONTH, levels=month.name), y=`Regional ID`,  fill=catchByRegionID, colour=catchByRegionID,  
                  text=paste("Estimated # kept legal CN=", round(catchByRegionID,0), "\nBased on # confirmed biosamples, n=", nID)),
              stat="identity") +
    paletteer::scale_fill_paletteer_c("viridis::plasma", na.value = alpha("white", 0.1)) +
    paletteer::scale_colour_paletteer_c("viridis::plasma", na.value = alpha("gray80", 0.3)) +
    scale_x_discrete(labels=c("May", "Jun", "Jul", "Aug", "Sept")) +
    labs(x="", y="", fill="Estimated kept \nlegal CN catch \nby Regional ID \nArea 21", 
         colour="Estimated kept \nlegal CN catch \nby Regional ID \nArea 21") +
    facet_wrap(~YEAR, nrow=1) + 
    theme_bw() +
    theme(axis.title = element_text(face="bold", size=10),
          axis.text.y = element_text(colour="black", size=8),
          axis.text.x = element_text(colour="black", size=9, angle=45, hjust=1),
          legend.title = element_text(face="bold", size=12),
          legend.text = element_text(size=8),
          #legend.key.size = unit(5, "mm"),
          #legend.position = "bottom",
          #legend.direction = "horizontal",
          strip.text = element_text(size=10))
  ,
  tooltip = c("MONTH", "Regional ID", "text")) 
```

<br>

<br>

```{r a121comp, out.width='1300px', out.height='500px', fig.cap='Estimated kept, legal (<45cm, >80cm) Chinook catch in Area 121. ID is to the Region-level in all cases except for Nitinat River. Estimates calculated by applying the proportion of Nitinat chinook detected in creel biosamples to total year-month-area catch estimates. **NOTE 2023 and 2024 samples in incomplete as of January 2025.**'}

# AREA 121

plotly::ggplotly(
  ggplot() +
    geom_tile(data=WCVIcrestBDWR.catEst %>% 
                filter(AREA=="121") %>% 
                group_by(YEAR, MONTH, AREA, NitRegionID) %>% 
                summarize(nID = sum(n_origin, na.rm=T),
                          TotalCatch = unique(Est_Kept)) %>% 
                group_by(YEAR, MONTH, AREA) %>% 
                mutate(totalSamples = sum(nID),
                       propn_NitRegionID = nID/totalSamples,
                       catchByRegionID = round(propn_NitRegionID*TotalCatch,0)) %>% 
                ungroup() %>%
                filter(!is.na(NitRegionID)) %>%
                full_join(data.frame(expand.grid(NitRegionID=WCVIcrestBDWR.catEst %>% 
                                                   filter(AREA=="121") %>% 
                                                   group_by(NitRegionID) %>% 
                                                   summarize(NitRegionID=unique(NitRegionID)) %>% 
                                                   filter(!is.na(NitRegionID)) %>%
                                                   pull(NitRegionID),
                                                 YEAR = c(min(WCVIcrestBDWR.catEst$YEAR):max(WCVIcrestBDWR.catEst$YEAR)),
                                                 MONTH = c("May","June","July","August","September"),
                                                 AREA = "121")),
                          .,
                          by=c("YEAR", "MONTH", "AREA", "NitRegionID")) %>%
                complete(., YEAR, MONTH, AREA, NitRegionID) %>%
                filter(!is.na(NitRegionID)) %>%
                rename(`Regional ID`=NitRegionID)
              ,
              aes(x=factor(MONTH, levels=month.name), y=`Regional ID`,  fill=catchByRegionID, colour=catchByRegionID,  
                  text=paste("Estimated # kept legal CN=", round(catchByRegionID,0), "\nBased on # confirmed biosamples, n=", nID)),
              stat="identity") +
    paletteer::scale_fill_paletteer_c("viridis::plasma", na.value = alpha("white", 0.1)) +
    paletteer::scale_colour_paletteer_c("viridis::plasma", na.value = alpha("gray80", 0.3)) +
    scale_x_discrete(labels=c("May", "Jun", "Jul", "Aug", "Sept")) +
    labs(x="", y="", fill="Estimated kept \nlegal CN catch \nby Regional ID \nArea 121", 
         colour="Estimated kept \nlegal CN catch \nby Regional ID \nArea 121") +
    facet_wrap(~YEAR, nrow=1) + 
    theme_bw() +
    theme(axis.title = element_text(face="bold", size=10),
          axis.text.y = element_text(colour="black", size=8),
          axis.text.x = element_text(colour="black", size=9, angle=45, hjust=1),
          legend.title = element_text(face="bold", size=12),
          legend.text = element_text(size=8),
          #legend.key.size = unit(5, "mm"),
          #legend.position = "bottom",
          #legend.direction = "horizontal",
          strip.text = element_text(size=10))
  ,
  tooltip = c("MONTH", "Regional ID", "text")) 
```

<br>


<br>

## **Sample rate**

Sample rate is simply the number of samples collected divided by the estimated kept catch, calculated for each month/area/year. It can give an indication of sampling effort to interpret catch trends. 

```{r include=F}
samp_Rate <- left_join(catEst %>% 
                         filter(DISPOSITION=="Kept", SPECIES=="CHINOOK SALMON") %>%
                         group_by(YEAR, MONTH, PFMA) %>%
                         summarize(TotalCatch = sum(ESTIMATE, na.rm=T)) %>%
                         arrange(YEAR, PFMA, match(MONTH, month.name)) %>%
                         mutate(PFMA = str_sub(PFMA, 6,9)) ,
                       WCVIcrestBDWR %>% 
                         filter(SAMPLE_TYPE=="Sport", SPECIES==124, DISPOSITION=="Kept", LEGAL_STATUS=="LEGAL") %>% 
                         group_by(YEAR, MONTH, AREA) %>% 
                         summarize(n = n()) ,
                       by=c("PFMA"="AREA", "YEAR", "MONTH")) %>%
  mutate(SampleRate = round(n/TotalCatch,3)) %>%
  print()
```


```{r sampRateFig, out.width='1500px', out.height='600px', fig.cap='Sample rate or "sampling effort" of kept legal Chinook catch in all WCVI areas from 2017-2024. Sampling rate or effort is defined as the number of legal-sized biological samples collected in each area-month-year divded by the total kept legal catch estimate for that area-month-year. Catch estimates are point estimates; this does not take into account uncertainty around catch estimates.'}

ggplot(data=samp_Rate %>%
         filter(MONTH%in%c("May", "June", "July", "August", "September")), 
       aes(x=factor(MONTH, levels=month.name), y=SampleRate, group=as.factor(YEAR), colour=as.factor(YEAR), fill=as.factor(YEAR), 
           #text=paste0(MONTH, sep=" ", YEAR)
       )) +
  geom_line(size=0.5, alpha=0.7) +
  geom_point(size=2, alpha=0.7, shape=21) +
  scale_y_continuous(labels = scales::percent, limits = c(0,0.3)) +
  scale_x_discrete(labels=c("May","Jun","Jul","Aug","Sept")) +
  labs(x="", y="Sample Rate") +
  theme_bw() +
  theme(axis.text = element_text(size=8, colour="black"),
        axis.text.x = element_text(angle=45, hjust=1),
        axis.title = element_text(size=9, face="bold"),
        legend.background = element_rect(colour="black"),
        legend.position = c(0.85, 0.1),
        legend.text = element_text(size=7),
        legend.title = element_blank(),
        strip.text = element_text(size=9)) +
  facet_wrap(~PFMA, nrow=3, scales="free_y") +
  guides(fill = guide_legend(ncol = 2),
         colour = guide_legend(ncol = 2))
```

<br>








```{r include=F}
# ========================= DATA EXPORT =========================

readme <- data.frame("x1" = c("Last update: ",
                              "Created by: ",
                              "Source biodata from: ",
                              "Source biodata last download: ",
                              "Source catch data from: "),
                     "x2" = c(as.character(Sys.Date()),
                              Sys.info()[7][[1]],
                              "CREST 'Biological Data With Results'",
                              "January 2025",
                              "//ENT.dfo-mpo.ca/DFO-MPO/GROUP/PAC/Reg_Shares/FHM/SMCS/Salmon/FMCR_Fishery_Monitoring_Catch_Reporting/Recreational_CM/Catch_Data/SC Sport Catch Creel Sub-area Disposition (Master Do No Edit).xlsx"
                              )
                     )

# ==================== CREATE EXCEL FILE ==================== 

# Create Workbook ---------------------------
Nit.WCVIrec <- openxlsx::createWorkbook()

# Add tabs to Workbook ---------------------------
openxlsx::addWorksheet(Nit.WCVIrec, "readme")
openxlsx::addWorksheet(Nit.WCVIrec, "All Nitinat catch estimates")
openxlsx::addWorksheet(Nit.WCVIrec, "Nitinat Hat catch estimates")



# Add data to tabs ---------------------------
openxlsx::writeData(Nit.WCVIrec, sheet="readme", x=readme)
openxlsx::writeData(Nit.WCVIrec, sheet="All Nitinat catch estimates", 
                    x = WCVIcrestBDWR.catEst.Nit %>% 
                      group_by(YEAR, MONTH, AREA, isNit2) %>% 
                      summarize(`Stock ID2 n` = sum(n_lvl1, na.rm=T),
                                `Total Kept Legal Catch per Area-Month-Year` = unique(Est_Kept),
                                #`Est_Released Legal` = unique(`Est_Released Legal`)
                                )  %>%
                      group_by(YEAR, MONTH, AREA) %>% 
                      mutate(`Total Samples per Area-Month-Year` = sum(`Stock ID2 n`, na.rm=T),
                             `Stock ID2 % (n/Total)` = `Stock ID2 n`/`Total Samples per Area-Month-Year`,
                             `Estimated Kept Legal Catch by Stock ID2 (%*total catch)` = 
                               round(`Stock ID2 % (n/Total)`*`Total Kept Legal Catch per Area-Month-Year`, 0),
                             #NitOnly_forPlot = case_when(grepl("Other", isNit2) ~ Est_Kept-estKeptByID,
                            #                             TRUE ~ estKeptByID),
                            # NitOnly_forPlot = case_when(NitOnly_forPlot==0 ~ NA,
                             #                            TRUE ~ NitOnly_forPlot)
                            ) %>%
                      rename(`Stock ID2` = isNit2) %>%
                      ungroup() %>%
                      filter(!is.na(YEAR)) %>% 
                      full_join(data.frame(AREA=c("127", "27", "126", "26", "125", "25", "124", "24", "123", "23", "121", "22", 
                                                  "21", "20")) %>%
                                  mutate(YEAR=rep(2023, nrow(.)),
                                         MONTH=rep("May", nrow(.))),
                                .,
                                by=c("YEAR", "MONTH", "AREA")) %>%
                      complete(., YEAR, MONTH, AREA)
                    )

openxlsx::writeData(Nit.WCVIrec, sheet="Nitinat Hat catch estimates", 
                    x = WCVIcrestBDWR.catEst.Nit %>% 
                      group_by(YEAR, MONTH, AREA, isNit3) %>% 
                      summarize(`Stock ID3 n` = sum(n_lvl1, na.rm=T),
                                `Total Kept Legal Catch per Area-Month-Year` = unique(Est_Kept),
                                #`Est_Released Legal` = unique(`Est_Released Legal`)
                                )  %>%
                      group_by(YEAR, MONTH, AREA) %>% 
                      mutate(`Total Samples per Area-Month-Year` = sum(`Stock ID3 n`, na.rm=T),
                             `Stock ID3 % (n/Total)` = `Stock ID3 n`/`Total Samples per Area-Month-Year`,
                             `Estimated Kept Legal Catch by Stock ID3 (%*total catch)` = 
                               round(`Stock ID3 % (n/Total)`*`Total Kept Legal Catch per Area-Month-Year`, 0),
                             #NitOnly_forPlot = case_when(grepl("Other", isNit2) ~ Est_Kept-estKeptByID,
                            #                             TRUE ~ estKeptByID),
                            # NitOnly_forPlot = case_when(NitOnly_forPlot==0 ~ NA,
                             #                            TRUE ~ NitOnly_forPlot)
                            ) %>%
                      rename(`Stock ID3` = isNit3) %>%
                      ungroup() %>%
                      filter(!is.na(YEAR)) %>% 
                      full_join(data.frame(AREA=c("127", "27", "126", "26", "125", "25", "124", "24", "123", "23", "121", "22", 
                                                  "21", "20")) %>%
                                  mutate(YEAR=rep(2023, nrow(.)),
                                         MONTH=rep("May", nrow(.))),
                                .,
                                by=c("YEAR", "MONTH", "AREA")) %>%
                      complete(., YEAR, MONTH, AREA)
                    )



# ==================== EXPORT EXCEL FILE ==================== 

# To github ---------------------------
openxlsx::saveWorkbook(Nit.WCVIrec,
                       file=paste0(here::here("outputs"),
                                   "/R_OUT - WCVI Nitinat rec catch supplementary data ",
                                   min(WCVIcrestBDWR.catEst.Nit$YEAR, na.rm=T),
                                   "-",
                                   max(WCVIcrestBDWR.catEst.Nit$YEAR, na.rm=T),
                                   ".xlsx"),
                       overwrite=T,
                       returnValue=T)






```

