

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning=F, message=F)

#library(here)
library(tidyverse)
library(readxl)
library(kableExtra)
library(bookdown)

"%notin%" <- Negate("%in%")

#source("C:/Users/DAVIDSONKA/Documents/ANALYSIS/San_Juan/san_juan_CN/scripts/functions/getExtractorData.R", local = knitr::knit_global())
source("C:/Users/DAVIDSONKA/Documents/ANALYSIS/San_Juan/san_juan_CN/scripts/functions/getMrpPadsData.R", local = knitr::knit_global())
source("C:/Users/DAVIDSONKA/Documents/ANALYSIS/San_Juan/san_juan_CN/scripts/functions/getNusedsData.R", local = knitr::knit_global())


# ====================== ENUMERATION data ====================== 
nuseds_yrs <- getNuSEDS("C:/Users/DAVIDSONKA/Documents/ANALYSIS/San_Juan/san_juan_CN/scripts/json/nuseds_esc_query_SJ.json", password=NULL)%>%
                  filter(!grepl("SOUTH THOMPSON", `Conservation Unit Name`), `Waterbody Name`=="SAN JUAN RIVER", Species=="Chinook") %>%
                  arrange(`Analysis Year`) %>%
                  pull(`Analysis Year`)

    
# Long-term time series (merge NuSEDS and Esc Index for complete time series) ----------------------- 
sj.esc <- full_join(
  #here() won't work unfortunately --\/
  getNuSEDS("C:/Users/DAVIDSONKA/Documents/ANALYSIS/San_Juan/san_juan_CN/scripts/json/nuseds_esc_query_SJ.json", password=NULL) %>%   
    filter(!grepl("SOUTH THOMPSON", `Conservation Unit Name`), `Waterbody Name`=="SAN JUAN RIVER", Species=="Chinook") %>%   # Remove the Harris Creek from S Thompson/Bessette 
    mutate(across(c(`Analysis Year`, `Max Estimate`:`Natural Adult Spawners`, `Other Adults Removals`:`Total Jack Return River`), as.numeric)) %>%
    mutate(`Total Adult Return River` = case_when(is.na(`Total Adult Return River`) ~ `Max Estimate`,
                                                  TRUE ~ `Total Adult Return River`),
           `Adult Broodstock Removals` = case_when(is.na(`Adult Broodstock Removals`) ~ `Total Broodstock Removals`,
                                                   TRUE ~ `Adult Broodstock Removals`)) %>% 
    pivot_longer(cols=c("Max Estimate":"Unspecified Return", "Other Removals":"Natural Adult Spawners", "Other Adults Removals":"Total Jack Return River"), 
                 names_to = "est_type", values_to = "estimate") %>% 
    rename(waterbody_name=`Waterbody Name`,
           analysis_year=`Analysis Year`) %>% 
    mutate(source="NuSEDS"),
  
  read_excel("//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/WCVI/ESCAPEMENT/Data/NEW ESCAPEMENT INDEX.xls", sheet="EscData", skip=4) %>%
    select(-c(`...17`)) %>% 
    filter(Year %notin% nuseds_yrs) %>%
    mutate(across(c("Ck Escape":"Brd Rem Cm"), ~ifelse(. %in% c("AP", "NO", "t"), NA, as.numeric(.)))) %>%
    select(c(Area:`Brd Rem Ck`)) %>%
    pivot_longer(cols=c("Ck Escape","Brd Rem Ck"), names_to = "est_type", values_to="estimate") %>%
    mutate(Species="Chinook") %>%
    mutate(est_type = case_when(grepl("Brd Rem", est_type) ~ "Adult Broodstock Removals",
                                grepl("Esc", est_type) ~ "Natural Adult Spawners")) %>%     
    #filter(Year%in%c((as.numeric(max(sj.esc.raw$`Analysis Year`))+1):lubridate::year(Sys.Date())), System=="San Juan") %>%
    filter(Year%in%c(2021,2022,2004), System=="San Juan") %>%
    rename(analysis_year=Year,
           waterbody_name=System) %>% 
    mutate(source="New Esc Index") %>%
    mutate_at("Area", as.character)
) %>% 
  filter() %>%
  print()



# Timing -----------------------
sj.cn.timing <- read_excel("//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/WCVI/ESCAPEMENT/Data/run timing tables (updated).xlsx", sheet="San Juan", range="A4:W65") %>% 
  rename(year=`...1`,
         system=`81`) %>%
  fill(year) %>%
  filter(year!=max(year) | !is.na(system)) %>% 
  pivot_longer(`82`:`124`, names_to = "stat_week", values_to = "count") %>%
  filter(system=="SJ", !is.na(stat_week))
sj.cn.timing$stat_week <- factor(sj.cn.timing$stat_week, levels=c("81","82","83","84","91","92","93","94","101","102","103","104","105",
                                                                  "111","112","113","114","115","121","122","123","124", ordered=T))




# Hatchery releases for comparison -----------------------
sj.releases.sep <- read_excel("C:/Users/DAVIDSONKA/Documents/ANALYSIS/San_Juan/data/escapement/SEP RELEASES_all-years-spp_2022-10-31.xlsx", sheet="Actual Release") %>%
  filter(FACILITY_NAME=="San Juan River H", SPECIES_NAME=="Chinook") 



# ====================== AGE data ====================== 
sj.FULL.age <- full_join(
  #sj.NuSEDS.age <- 
    getNuSEDS("C:/Users/DAVIDSONKA/Documents/ANALYSIS/San_Juan/san_juan_CN/scripts/json/nuseds_padsCU_query_SJ.json", 
                             password=NULL) %>%
      mutate(data_source = "NuSEDS") %>%
      rename(ScaleCondition = `Part Age Code`,
             FishEdge = `Edge Code`,
             AgeingStructure = `Reading Method`,
             FormatName = Format,
             GrAge = `GR Age`,
             EuAge = `EU Age`,
             FishNumber = `Container Address`,
             ProjectName = Project,
             RecoveryYear = `Fiscal Year`,
             GearMrpName.y = Gear,
             LifeHistory = `Life History Stage`,
             CntStartDate = `Sample Start Date`,
             CntEndDate = `Sample End Date`,
             ContainerId = `Container Label`) %>%
      mutate_at("RecoveryYear", as.integer) %>%
      mutate_at(c("CntStartDate", "CntEndDate"), as.Date),
  
  #sj.PADS.age <- 
    allPADS %>%
      filter(Location=="San Juan River", Species=="Chinook") %>%
      mutate(data_source = "MRP") %>%
      mutate_at("FishNumber", as.character)
) %>%
  mutate(Species = str_to_title(Species),
         Location = str_to_title(Location),
         FormatName = str_to_lower(FormatName),
         `(R) SCALE BOOK NUM` = ContainerId,
         `(R) SCALE CELL NUM` = FishNumber,
         `(R) SCALE BOOK-CELL CONCAT` = case_when(!is.na(ContainerId) & !is.na(FishNumber) ~ paste0(ContainerId,sep="-",FishNumber)),
         `(R) RESOLVED AGE` = case_when(GrAge %in% c(10,11) ~ 1,
                                        GrAge %in% c(20,21,22) ~ 2,
                                        GrAge %in% c(30,21,32,33) ~ 3,
                                        GrAge %in% c(40,41,42,43,44) ~ 4,
                                        GrAge %in% c(50,51,52,53,54,55) ~ 5,
                                        GrAge %in% c(60,61,62,63,64,65,66) ~ 6)) %>%
  filter(grepl("Chinook", Species)) %>%
  print()





# ==============================
# Age comp each year
sj_age_summary <- sj.FULL.age %>%
  filter(!is.na(`(R) RESOLVED AGE`)) %>%
  group_by(RecoveryYear, `(R) RESOLVED AGE`) %>% 
  summarize(count = n()) %>%
  group_by(RecoveryYear) %>% 
  mutate(year_total = sum(count),
         propn = count/year_total) %>%
  print()

# Long-term average age comp
sj_LT_avgs <- sj_age_summary %>%
  group_by(`(R) RESOLVED AGE`) %>% 
  summarize(avg_propn = round(mean(propn)*100,1),
            sd_propn = round(sd(propn)*100,1)) %>%
  rename(resolved_age = `(R) RESOLVED AGE`) %>%
  arrange(desc(avg_propn)) %>%
  slice(1:2) %>%
  print()

# Recent-year average age comp
sj_RY_avgs <- sj_age_summary %>%
  filter(RecoveryYear%in%c(lubridate::year(Sys.Date())-10) : lubridate::year(Sys.Date())) %>%
  group_by(`(R) RESOLVED AGE`) %>% 
  summarize(avg_propn = round(mean(propn)*100,1),
            sd_propn = round(sd(propn)*100,1)) %>%
  rename(resolved_age = `(R) RESOLVED AGE`) %>%
  arrange(desc(avg_propn)) %>%
  slice(1:2) %>%
  print()





### *******NEXT DAY:
# -- fix references (thjink need to add bookdown pkg)
## ---fix fig caps
## --- transfer to child rmd,  try rendering parent script
## ---- add method field to sj.cn.timing file (get from nuseds - extract and join? ?)
```


# **San Juan Chinook escapement: abundance and biodata** {#esc}

Status snapshot: 

* Chinook enhancement began ~1975 

* Recent 10-yr average adult escapement (incl broodstock): `r sj.esc %>% filter(est_type=="Total Adult Return River") %>% arrange(desc(analysis_year)) %>% slice(1:10) %>% summarize(mean=round(mean(estimate,na.rm=T),0)) %>% pull(mean)`

  + Range: `r sj.esc %>% filter(est_type=="Total Adult Return River") %>% arrange(desc(analysis_year)) %>% slice(1:10) %>% summarize(min=min(estimate)) %>% pull(min)` - `r sj.esc %>% filter(est_type=="Total Adult Return River") %>% arrange(desc(analysis_year)) %>% slice(1:10) %>% summarize(max=max(estimate)) %>% pull(max)`

* Recent 4-yr avg adult escapement (incl broodstock): `r sj.esc %>% filter(est_type=="Total Adult Return River") %>% arrange(desc(analysis_year)) %>% slice(1:4) %>% summarize(mean=round(mean(estimate,na.rm=T),0)) %>% pull(mean)`

  + Range: `r sj.esc %>% filter(est_type=="Total Adult Return River") %>% arrange(desc(analysis_year)) %>% slice(1:4) %>% summarize(min=min(estimate)) %>% pull(min)` - `r sj.esc %>% filter(est_type=="Total Adult Return River") %>% arrange(desc(analysis_year)) %>% slice(1:4) %>% summarize(max=max(estimate)) %>% pull(max)`

* Long-term dominant age structure (1990-present): `r as.list(sj_LT_avgs[1,1])` (~`r as.list(sj_LT_avgs[1,2])`%),  `r as.list(sj_LT_avgs[2,1])` (~`r as.list(sj_LT_avgs[2,2])`%) year olds

* Recent (10yr) dominant age structure: `r as.list(sj_RY_avgs[1,1])` (~`r as.list(sj_RY_avgs[1,2])`%),  `r as.list(sj_RY_avgs[2,1])` (~`r as.list(sj_RY_avgs[2,2])`%) year olds

* Hatchery vs wild: uncertain due to historical marking/tagging process. Mostly unclipped with no thermal mark (but historical TM process highly unreliable)


<br>

## **Escapement and run timing** {#abundance}

San Juan Chinook escapement has varied over time. From its peak at 7500 in 1970 it declined to less than 500 fish between 1975-1978, but has since returned to escapements that tend to oscillate around 2000-4000 fish (see Figure \@ref(fig:sj-esc-fig)). While broodstock collections are only documented in NuSEDS from 1995 onward, releases began in the 70's; see Enhancement Activity information (section \@ref(enhancement)) below. 

```{r sj-esc-fig, fig.cap='Spawners and broodstock removals for San Juan Chinook, specific adults only where known. Vertical dashed lines indicate periods of enhancement. Long-term averages (solid gray line) are calculated across the whole time series since enhancement began. Recent averages (black line) are calculated from the most recent 4 years of complete data. Data from NuSEDS, excluding tributaries. Red outlined bars (if any) indicate escapement estimates not in NuSEDS.'}
# FIGURE: All species ----------------------- 
ggplot() +
  # Enhancement start vertical line:
  geom_vline(data=sj.releases.sep%>%group_by(SPECIES_NAME)%>%summarize(start=min(BROOD_YEAR))%>%rename(Species=SPECIES_NAME)%>%
               mutate(n=7000),
             aes(xintercept=start), colour="gray60", linetype="dashed", size=0.5) +
  # Escapement goal and average escapements: 
  geom_hline(data=data.frame(Species="Chinook", Smsy=2200, Sgen=990, Srep=5200), aes(yintercept=Smsy), size=1, linetype="dashed") +
  geom_hline(data=sj.esc %>% filter(analysis_year>=(max(analysis_year-10)), est_type=="Natural Adult Spawners") %>% 
               summarize(mean=round(mean(n,na.rm=T),0)),
             aes(yintercept=mean), size=1, linetype="solid", colour="black") +
  # Plot NuSEDS escapement data:
  geom_bar(data=sj.esc %>% filter(source=="NuSEDS", est_type %in% c("Unspecified Return", "Natural Adult Spawners", "Adult Broodstock Removals"),
                                  !is.na(estimate)),
           aes(x=analysis_year, y=as.numeric(estimate), fill=est_type, colour=est_type), stat="identity", alpha=0.8, width=0.9, size=0.3) +
  # Add in any prelim estimates from NEW ESCAPEMENT INDEX (depends on time of year):
  geom_bar(data=sj.esc%>%filter(est_type%in%c("Natural Adult Spawners", "Adult Broodstock Removals"), source=="New Esc Index"), 
           aes(x=analysis_year, y=estimate, fill=est_type, colour=est_type), stat="identity", alpha=0.7, width=0.9, size=0.5) +
  
  # Highlight the prelim NEW ESCAPEMENT INDEX estimates in red to indicate preliminary data: 
  geom_bar(data=sj.esc%>%filter(est_type%in%c("Natural Adult Spawners", "Adult Broodstock Removals"), source=="New Esc Index")%>%
             group_by(analysis_year,Species)%>%summarize(total=sum(estimate)), 
           aes(x=analysis_year, y=total), stat="identity", width=0.9, size=0.7, fill="transparent", colour="red") +
  
  scale_x_continuous(breaks=seq(min(sj.esc$analysis_year), max(sj.esc$analysis_year), by=5)) +
  scale_fill_manual(breaks = c("Natural Adult Spawners", "Adult Broodstock Removals", "Unspecified Return"), 
                    values=c("dodger blue","orange", "gray")) +
  scale_colour_manual(breaks = c("Natural Adult Spawners", "Adult Broodstock Removals", "Unspecified Return"), 
                      values=c("dodger blue","orange", "gray")) +
  labs(x="Return Year", y="Number of Chinook") +
  theme_bw() +
  theme(axis.text = element_text(colour="black",size=11),
        axis.text.x = element_text(angle=45,hjust=1),
        axis.title = element_text(face="bold",size=13),
        panel.grid.major.y = element_line(colour="gray70"),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size=10),
        legend.position = c(0.75,0.85),
        #legend.direction = "horizontal",
        legend.background = element_blank(),
        strip.text = element_text(size=10)) +
  guides(colour = guide_legend(override.aes=list(colour=NA))) 
```




