---
title: "part1-esc-knitable"
author: "South Coast StA"
date: 'Last update: `r format(Sys.time(), sep="_", "%Y-%m-%d %H:%M")`'
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_collapsed: false 
    toc_depth: 5
    number_sections: true
params:
    password:
      label: "Enter computer password"
      value: ""
      input: password
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning=F, message=F)

#library(here)
library(tidyverse)
library(readxl)
library(kableExtra)
library(bookdown)

"%notin%" <- Negate("%in%")

#source("C:/Users/DAVIDSONKA/Documents/ANALYSIS/San_Juan/san_juan_CN/scripts/functions/getExtractorData.R", local = knitr::knit_global())
source("C:/Users/DAVIDSONKA/Documents/ANALYSIS/San_Juan/san_juan_CN/scripts/functions/getMrpPadsData.R", local = knitr::knit_global())
source("C:/Users/DAVIDSONKA/Documents/ANALYSIS/San_Juan/san_juan_CN/scripts/functions/getNusedsData.R", local = knitr::knit_global())


# ====================== ENUMERATION data ====================== 

a21_22.esc <- getNuSEDS("C:/Users/DAVIDSONKA/Documents/ANALYSIS/Area21-22/A21-22/scripts/json/nuseds_esc_query_A21-22.json", password=NULL) %>% 
    mutate(across(c(`Analysis Year`, `Max Estimate`:`Natural Adult Spawners`, 
                    `Other Adults Removals`:`Total Jack Return River`), as.numeric)) %>%
  print()


# BY escapements ------------------------
adultesc <- a21_22.esc %>% 
  rowwise() %>%
  mutate(`Total Adult Return River` = case_when(is.na(`Total Adult Return River`) ~ 
                                                  sum(`Natural Adult Spawners`,`Adult Broodstock Removals`,na.rm=T),
                                                TRUE~`Total Adult Return River`) ) %>%
  group_by(Area, `Analysis Year`, Species, `Waterbody Name`, `Estimate Classification`) %>%
  summarize(total_adults=sum(`Total Adult Return River`, na.rm=T), SENPA = unique(`SEN Presence Adult`)) %>% 
  #mutate(`Waterbody Name` = paste0(`Waterbody Name`, sep=" (", Area, sep=")")) %>%
  print()

by.esc <- 
  full_join(
    adultesc %>% 
      filter(Species=="Chinook" & `Analysis Year`%in%c(lubridate::year(Sys.Date())-5, lubridate::year(Sys.Date())-4, lubridate::year(Sys.Date())-3)), 

    adultesc %>%
      filter(Species=="Chum" & `Analysis Year`%in%c(lubridate::year(Sys.Date())-4, lubridate::year(Sys.Date())-3)) 
  ) %>%
  full_join(.,
            adultesc %>% 
              filter(Species=="Coho" & `Analysis Year`%in%c(lubridate::year(Sys.Date())-4, lubridate::year(Sys.Date())-3)) 
            ) %>%
  full_join(.,
            adultesc %>% 
              filter(Species=="Sockeye" & `Analysis Year`%in%c(lubridate::year(Sys.Date())-4, lubridate::year(Sys.Date())-5))
            ) %>%
  full_join(.,
            adultesc %>% 
              filter(Species=="Pink" & `Analysis Year`%in%c(lubridate::year(Sys.Date())-2)) 
            ) %>%
  arrange(Area, `Waterbody Name`, Species, `Analysis Year`) %>%
  select(Area, `Waterbody Name`, Species, `Analysis Year`, total_adults, `Estimate Classification`, SENPA)
write.csv(by.esc, "C:/Users/DAVIDSONKA/Documents/ANALYSIS/Area21-22/outputs/area21-22_2023return_BYesc.csv" , row.names=F)




# ====================== TIMING data ====================== 

nit.timing <- full_join(
  read_excel("//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/WCVI/ESCAPEMENT/Data/run timing tables (updated).xlsx", sheet="Nitinat", range="A4:W32") %>% 
    rename(year=`...1`) %>%
    mutate(`92` = case_when(grepl("-", `92`) ~ 1500, TRUE~as.numeric(`92`)),
           species="Chinook") %>%
    pivot_longer(`81`:`124`, names_to = "stat_week", values_to = "count"),
  
  read_excel("//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/WCVI/ESCAPEMENT/Data/run timing tables (updated).xlsx", sheet="Nitinat", range="A38:W66") %>% 
    rename(year=`...1`) %>%
    mutate(`92` = case_when(grepl("-", `92`) ~ 1500, TRUE~as.numeric(`92`)),
           species="Coho") %>%
    pivot_longer(`81`:`124`, names_to = "stat_week", values_to = "count")
) %>% 
  full_join(.,
            read_excel("//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/WCVI/ESCAPEMENT/Data/run timing tables (updated).xlsx", sheet="Nitinat", range="A72:W100") %>% 
              rename(year=`...1`) %>%
              mutate(`92` = case_when(grepl("-", `92`) ~ 1500, TRUE~as.numeric(`92`)),
                     species="Chum") %>%
              pivot_longer(`81`:`124`, names_to = "stat_week", values_to = "count")
            ) %>%
  print()
nit.timing$stat_week <- factor(nit.timing$stat_week, levels=c("81","82","83","84","91","92","93","94","101","102","103","104","105",
                                                              "111","112","113","114","115","121","122","123","124", ordered=T))




# ====================== (minimal) HATCHERY data ====================== 

# Hatchery releases for comparison -----------------------
#sj.releases.sep <- read_excel("C:/Users/DAVIDSONKA/Documents/ANALYSIS/San_Juan/data/enhancement-PNI/SEP #RELEASES_all-years-spp_2022-10-31.xlsx", sheet="Actual Release") %>%
#  filter(FACILITY_NAME=="San Juan River H", SPECIES_NAME=="Chinook") 




# ====================== AGE data ====================== 

a21_22.age <- full_join(
  a21_22.NuSEDS.age <- 
    getNuSEDS("C:/Users/DAVIDSONKA/Documents/ANALYSIS/Area21-22/A21-22/scripts/json/nuseds_padsCU_query_A21-22.json", password=NULL) %>%
      mutate(data_source = "NuSEDS",
             Species = gsub('.{7}$', '', Species)) %>%
      rename(ScaleCondition = `Part Age Code`,
             FishEdge = `Edge Code`,
             AgeingStructure = `Reading Method`,
             FormatName = Format,
             GrAge = `GR Age`,
             EuAge = `EU Age`,
             FishNumber = `Container Address`,
             ProjectName = Project,
             RecoveryYear = `Fiscal Year`,
             GearMrpName.y = Gear,
             LifeHistory = `Life History Stage`,
             CntStartDate = `Sample Start Date`,
             CntEndDate = `Sample End Date`,
             ContainerId = `Container Label`) %>%
      mutate_at("RecoveryYear", as.integer) %>%
      mutate_at(c("CntStartDate", "CntEndDate"), as.Date),
  
  a21_22.PADS.age <- 
    allPADS %>%
      filter(Area %in% c(21, 22), grepl("Escapement", ProjectName)) %>%
      mutate(data_source = "MRP") %>%
      mutate_at("FishNumber", as.character)
) %>%
  mutate(Species = str_to_title(Species),
         Location = str_to_title(Location),
         FormatName = str_to_lower(FormatName),
         `(R) SCALE BOOK NUM` = ContainerId,
         `(R) SCALE CELL NUM` = FishNumber,
         `(R) SCALE BOOK-CELL CONCAT` = case_when(!is.na(ContainerId) & !is.na(FishNumber) ~ paste0(ContainerId,sep="-",FishNumber)),
         `(R) RESOLVED AGE` = case_when(GrAge %in% c(10,11) ~ 1,
                                        GrAge %in% c(20,21,22) ~ 2,
                                        GrAge %in% c(30,21,32,33) ~ 3,
                                        GrAge %in% c(40,41,42,43,44) ~ 4,
                                        GrAge %in% c(50,51,52,53,54,55) ~ 5,
                                        GrAge %in% c(60,61,62,63,64,65,66) ~ 6)) %>%
  filter(LifeHistory == "Adult") %>%
  print()



# ============================== BIODATA ==============================

#sj.biodata <- read_excel("//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/SC_BioData_Management/2-Escapement/2015-2023_WCVI_Escapement-FSC_BioData.xlsx", sheet="Biodata 2015-2022", guess_max=20000)
#sj.fecund <- read_excel("//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/SC_BioData_Management/2-Escapement/2015-2023_WCVI_Escapement-FSC_BioData.xlsx", sheet="SJFecundity", skip=1)
#sj.SR <- read_excel("//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/SC_BioData_Management/2-Escapement/2015-2023_WCVI_Escapement-FSC_BioData.xlsx", sheet="SJsexRatio", skip=1)





# ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~



```

# **Nitinat escapement: abundance and biodata** {#esc}

Status snapshot:

-   Chinook enhancement began \~1975

-   Recent 10-yr average adult escapement (incl broodstock):  

    -   Range:  

-   Recent 4-yr avg adult escapement (incl broodstock):  

    -   Range:  

-   Long-term dominant age structure (1990-present):  

-   Recent (10yr) dominant age structure:  

-   Hatchery vs wild:  

<br>

## **Escapement**

h (see Figure \@ref(fig:a21-fig))

```{r a21-fig, fig.cap='Spawner estimates for salmon in Area 21. Vertical dashed lines indicate periods of enhancement. Long-term averages (solid gray line) are calculated across the whole time series since enhancement began. Recent averages (black line) are calculated from the most recent 4 years of complete data. Data from NuSEDS.'}
# FIGURE: All populations Area 21 ----------------------- 
ggplot() +
  # --- Enhancement start vertical line:
  #geom_vline(data=sj.releases.sep%>%group_by(SPECIES_NAME)%>%summarize(start=min(BROOD_YEAR))%>%rename(Species=SPECIES_NAME)%>%
  #             mutate(n=7000),
  #           aes(xintercept=start), colour="gray60", linetype="dashed", size=0.5) +
  # --- Escapement goal and average escapements: 
  #geom_hline(data=data.frame(Species="Chinook", Smsy=2200, Sgen=990, Srep=5200), aes(yintercept=Smsy), size=1, linetype="dashed") +
  #geom_hline(data=sj.esc %>% filter(analysis_year>=(max(analysis_year-10)), est_type=="Natural Adult Spawners") %>% 
  #             summarize(mean=round(mean(n,na.rm=T),0)),
  #           aes(yintercept=mean), size=1, linetype="solid", colour="black") +
  # --- Plot NuSEDS escapement data:
  geom_bar(data=a21_22.esc %>% filter(Area==21, est_type %in% c("Unspecified Return", "Natural Adult Spawners", "Adult Broodstock Removals"),
                                  !is.na(estimate)),
           aes(x=analysis_year, y=as.numeric(estimate), fill=est_type, colour=est_type), stat="identity", alpha=0.8, width=0.9, size=0.3) +
  scale_x_continuous(breaks=seq(min(a21_22.esc$analysis_year), max(a21_22.esc$analysis_year), by=5)) +
  scale_fill_manual(breaks = c("Natural Adult Spawners", "Adult Broodstock Removals", "Unspecified Return"), 
                    values=c("dodger blue","orange", "gray")) +
  scale_colour_manual(breaks = c("Natural Adult Spawners", "Adult Broodstock Removals", "Unspecified Return"), 
                      values=c("dodger blue","orange", "gray")) +
  labs(x="Return Year", y="Number of Fish") +
  theme_bw() +
  theme(axis.text = element_text(colour="black",size=11),
        axis.text.x = element_text(angle=45,hjust=1),
        axis.title = element_text(face="bold",size=13),
        panel.grid.major.y = element_line(colour="gray70"),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size=10),
        legend.position = c(0.7,0.1),
        #legend.direction = "horizontal",
        legend.background = element_blank(),
        strip.text = element_text(size=10)) +
  guides(colour = guide_legend(override.aes=list(colour=NA))) +
  facet_wrap(~Population, scales="free_y")
```

<br>

```{r a22-fig, fig.cap='Spawners estimates for salmon in Area 22. Vertical dashed lines indicate periods of enhancement. Long-term averages (solid gray line) are calculated across the whole time series since enhancement began. Recent averages (black line) are calculated from the most recent 4 years of complete data. Data from NuSEDS.'}
ggplot() +
  # --- Plot NuSEDS escapement data:
  geom_bar(data=a21_22.esc %>% filter(Area==22, est_type %in% c("Unspecified Return", "Natural Adult Spawners", "Adult Broodstock Removals"),
                                  !is.na(estimate)),
           aes(x=analysis_year, y=as.numeric(estimate), fill=est_type, colour=est_type), stat="identity", alpha=0.8, width=0.9, size=0.3) +
  scale_x_continuous(breaks=seq(min(a21_22.esc$analysis_year), max(a21_22.esc$analysis_year), by=7)) +
  scale_fill_manual(breaks = c("Natural Adult Spawners", "Adult Broodstock Removals", "Unspecified Return"), 
                    values=c("dodger blue","orange", "gray")) +
  scale_colour_manual(breaks = c("Natural Adult Spawners", "Adult Broodstock Removals", "Unspecified Return"), 
                      values=c("dodger blue","orange", "gray")) +
  labs(x="Return Year", y="Number of Fish") +
  theme_bw() +
  theme(axis.text = element_text(colour="black",size=11),
        axis.text.x = element_text(angle=45,hjust=1),
        axis.title = element_text(face="bold",size=13),
        panel.grid.major.y = element_line(colour="gray70"),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size=10),
        legend.position = c(0.7,0.1),
        #legend.direction = "horizontal",
        legend.background = element_blank(),
        strip.text = element_text(size=10)) +
  guides(colour = guide_legend(override.aes=list(colour=NA))) +
  facet_wrap(~Population, scales="free_y")
```


### **Escapement goals** {#esc-goals}

Current escapement goal developed by Holt et al. based on a habitat model (Parken et al 2008) using 356.62 sq-km watershed area:

-   Smsy: 2,400 (1,300-4,500) - The escapement required to produce maximum sustainable yield

-   Sgen: 1,100 (270-3,300) - The escapement required to recover to Smsy in 1 generation

-   Srep: 5,600 (3,100-10,000) - System capacity

-   Fish Stock Provisions Rebuilding goal = Sgen \* \~1.5 = `r 1100*1.5`

<br>

<br>

## **Run timing** {#timing}

Most chinook are observed in the San Juan between the third week of September (statweek 93) and the second week of October (statweek 102) (Figure \@ref(fig:nit-timing)). Most counts come from a combination of the fence (which is usually in until first or second week of October) and swim surveys (often later, but not always). The years that are swim survey only are .

```{r nit-timing, fig.cap='Weekly swim survey counts in the Nitinat River from 1995-2022 inclusive. Counts are more representative of arrival rather than peak spawning (no documentation of behaviour accompanies counts). DFO South Coast Stock Assessment data.'}
ggplot(nit.timing %>% mutate(year_group=case_when(year<2000~"pre 2000",
                                                    year%in%c(2000:2010) ~ "2000-2010",
                                                    year>2010~"post-2010"),
                               year_group = factor(year_group, levels=c("pre 2000", "2000-2010", "post-2010",ordered=T))), 
       aes(x=stat_week, y=count, group=year, fill=year_group, colour=year_group)) +
  geom_bar(stat="identity", alpha=0.7) +
  labs(x="Stat week", y="Count") +
  theme_bw() +
  theme(axis.text = element_text(colour="black"),
        axis.title = element_text(face="bold"),
        legend.title = element_blank()) +
  facet_wrap(~species, scales="free_y", nrow=3)
```

<br>

## **Upcoming return** {#forecast}

The `r lubridate::year(Sys.Date())` return will be composed of offspring from the following brood years:



```{r by-table}
by.esc %>% 
  mutate(across(c(total_adults), ~ format(.x, big.mark = ","))) %>% 
  rename(`Return year` = `Analysis Year`) %>%
  kbl(align="c", caption="BY escapements (spawners+broodstock) contributing to the 2023 return. Data from NuSEDS.") %>%
  kable_paper("hover", full_width=T, position = "center") %>%
  collapse_rows(columns=c(1:3), valign="middle", target=1) 
```

- The `r lubridate::year(Sys.Date())` Nitinat Chinook forecast is for 26,000 (18,000-34,000; 67% age-4). 

- The `r lubridate::year(Sys.Date())` Nitinat Chum forecast p50 is 87,000 (p25=64,000; p75=122,000).

- The `r lubridate::year(Sys.Date())` Robertson Creek hatchery coho forecast is 'moderate', marine survival 7.1% (down 4% from 2022)

- The `r lubridate::year(Sys.Date())` Carnation Creek wild coho forecast is 'moderate', marine survival 2.5% (up 140% from 2022)




<br>

<br>

## **Age structure** {#escAge}

(see Table \@ref(tab:age-tab) and Figure \@ref(fig:age-fig)).

<br>

```{r age-tab, echo=F, warning=F, message=F}
# TABLE: Overall age structure -----------------------
a21_22.age %>% 
  filter(!is.na(`(R) RESOLVED AGE`)) %>%
  group_by(Species, RecoveryYear, `(R) RESOLVED AGE`) %>% 
  summarize(n=n()) %>% 
  group_by(Species, RecoveryYear) %>% 
  mutate(propn=n/sum(n)) %>%
  group_by(Species, `(R) RESOLVED AGE`) %>% 
  summarize(avg_propn = mean(propn), sd_propn=sd(propn)) %>% 
  mutate(avg_propn=round(avg_propn,3)*100, sd_propn=round(sd_propn,3)*100) %>%
  unite(col="Average proportion (± SD)", c(avg_propn, sd_propn), sep=" ± ") %>%
  rename(Age=`(R) RESOLVED AGE`) %>%
  kbl(align="c", caption="Age structure of San Juan Chinook (excludes tributaries). Data from PADS (NuSEDS and MRPIS).") %>%
  kable_paper("hover", full_width=T, position = "center") 
```

<br>

```{r age-fig, echo=F, warning=F, message=F, fig.cap='Annual age structure of San Juan Chinook (excludes tributaries). Numbers indicate the number of samples contributing age data each year. Data from PADS (NuSEDS and MRPIS).'}
# FIGURE: AGE ~ TIME ----------------------- 
ggplot() +
  geom_bar(data=a21_22.age %>% 
         filter(!is.na(`(R) RESOLVED AGE`)) %>%
         group_by(Species, RecoveryYear, `(R) RESOLVED AGE`) %>% 
         summarize(n=n()) %>% 
         group_by(Species, RecoveryYear) %>% 
         mutate(sample_size = sum(n),
                propn=n/sum(n)),
       aes(x=RecoveryYear, y=propn, group=as.factor(`(R) RESOLVED AGE`), fill=as.factor(`(R) RESOLVED AGE`), colour=as.factor(`(R) RESOLVED AGE`)), 
       stat="identity", position="stack", width=0.5, alpha=0.8, size=1) +
  geom_text(data=a21_22.age %>% filter(!is.na(`(R) RESOLVED AGE`)) %>% group_by(Species, RecoveryYear) %>% summarize(sample_size = n()), 
            aes(x=RecoveryYear, y=1, label=sample_size), hjust=0, vjust=0, angle=45) +
  labs(y="Proportion", fill="Age:", colour="Age:") +
  scale_x_continuous(breaks=seq(1990,2050,by=2)) +
  scale_y_continuous(breaks=seq(0,1.1,by=0.25), limits=c(0,1.1)) +
  theme_bw() +
  theme(axis.text = element_text(colour="black", size=11),
        axis.text.x = element_text(angle=45, hjust=1),
        axis.title = element_text(face="bold", size=13),
        axis.title.x = element_blank(),
        legend.position = "bottom",
        legend.direction = "horizontal",
        legend.title = element_text(face="bold"),
        plot.caption = element_text(face="italic", hjust=0)) +
  facet_wrap(~Species)
```

```{r include=F}
# Long-term and recent age trends -----------------------

# Age comp each year
a21_22.age_summary <- a21_22.age %>%
  filter(!is.na(`(R) RESOLVED AGE`)) %>%
  group_by(Species, RecoveryYear, `(R) RESOLVED AGE`, GrAge) %>%
  summarize(Gr_n = n()) %>%
  group_by(Species, RecoveryYear, `(R) RESOLVED AGE`) %>%
  mutate(resolved_n = sum(Gr_n)) %>%
  group_by(Species, RecoveryYear) %>%
  mutate(year_n = sum(Gr_n),
         Rslv_propn = resolved_n/year_n,
         Gr_propn_of_rslv = Gr_n/resolved_n) %>%
  print()

# Long-term average age comp
a21_22.LT_avgs <- a21_22.age_summary %>%
  group_by(Species, `(R) RESOLVED AGE`, GrAge) %>%
  mutate(avg_Gr_propn_of_rslv = round(mean(Gr_propn_of_rslv)*100,1),
         sd_Gr_propn_of_rslv = round(sd(Gr_propn_of_rslv)*100,1)) %>%
  group_by(Species, `(R) RESOLVED AGE`) %>% 
  mutate(avg_rslv_propn = round(mean(Rslv_propn)*100,1),
         sd_rslv_propn = round(sd(Rslv_propn)*100,1)) %>%
  arrange(Species, desc(avg_rslv_propn)) %>%
  print()

a21_22.LT_avgs %>%
  group_by(Species, `(R) RESOLVED AGE`, GrAge) %>%
  summarize(avg_Gr_propn_of_rslv=unique(avg_Gr_propn_of_rslv),
            sd_Gr_propn_of_rslv = unique(sd_Gr_propn_of_rslv)) 

a21_22.LT_avgs %>%
  group_by(Species, `(R) RESOLVED AGE`) %>%
  summarize(avg_rslv_propn=unique(avg_rslv_propn),
            sd_rslv_propn = unique(sd_rslv_propn)) 

  
  
  
  
  
  

# Recent-year average age comp
a21_22.RY_avgs <- a21_22.age_summary %>%
  filter(RecoveryYear%in%c(lubridate::year(Sys.Date())-10) : lubridate::year(Sys.Date())) %>%
  group_by(Species, `(R) RESOLVED AGE`) %>%
  summarize(avg_propn = round(mean(propn)*100,1),
            sd_propn = round(sd(propn)*100,1)) %>%
  rename(resolved_age = `(R) RESOLVED AGE`) %>%
  arrange(desc(avg_propn)) %>%
  slice(1:2) %>%
  print()
  
```


<br>

<br>

## **Sex ratio and fecundity**

Very little sex ratio is available. We have two years of fence counts that may shed some light, although are likely highly uncertain.

```{r}
sj.SR %>% 
  filter(Species=="Chinook", Sex%in%c("M","F")) %>%
  group_by(Year, Sex) %>%
  summarize(n=sum(Count)) %>%
  group_by(Year) %>% 
  mutate(propn=round(n/sum(n)*100,1)) %>%
  rename(`Number counted by fence`=n, `Proportion of total (%)`=propn) %>% 
  kbl(align="c", caption="Sex ratio of chinook passing by the San Juan fence. Data from Four Mile Hatchery.") %>%
  #collapse_rows(columns=c(1,2), valign="middle", target=1) %>%
  kable_paper("hover", full_width=T, position = "center")
```

<br>

There is also very little fecundity data available, and all is bulk/annual averages from the hatchery.

```{r}
sj.fecund %>% 
  group_by(Year) %>%
  summarize(mean_fec = round(mean(Fecundity),0), sd=round(sd(Fecundity),0),
            n_females = sum(Number_females)) %>%
  mutate(across(c(mean_fec:n_females), ~ format(.x, big.mark = ","))) %>% 
  rename(`Mean fecundity`=mean_fec, `# females in sample`=n_females) %>%
  kbl(align="c", caption="Fecundity of San Juan Chinook with variance and sample size where known. Data from Four Mile Hatchery.") %>%
  #collapse_rows(columns=c(1,2), valign="middle", target=1) %>%
  kable_paper("hover", full_width=T, position = "center")
```

<br>

There is currently no estimate of natural female spawner success.
