---
title: "Nitinat escapement 2023 (PRELIM)"
author: "South Coast StA"
date: "Last update: `r Sys.Date()`"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_collapsed: false 
    toc_depth: 5
    number_sections: true
knit: (function(inputFile, encoding) { 
  rmarkdown::render(inputFile, encoding=encoding, output_file=paste0(substr(inputFile,1,nchar(inputFile)-4), "_PRELIMINARY_", format(Sys.Date(), 
  sep="_", "%Y-%m-%d"), '.html'))
  }
  )
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning=F, message=F)

# Function for appending date to html output - copy into YAML below output:
#

# Load libraries ----------------------------
library(tidyverse)
library(readxl)
library(here)
library(ggpubr)
library(kableExtra)


# Helpers ----------------------------
options(scipen = 9999999)
"%notin%" <- Negate("%in%")
analysis_year <- 2023


# ============================= Load data ============================
# OneDrive directory:
OD.dir <- paste("C:/Users", Sys.info()[6], "OneDrive - DFO-MPO/# Escapement Estimates/# 2023/Area 21-22/", sep="/")


# raw SILS ----------------------------
nit.sil <- readxl::read_excel("//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/WCVI/ESCAPEMENT/Data/2023/SILs/Area 21-22/Nitinat River/A22_Nitinat_River_2023.xlsx", sheet="Nitinat River raw SIL entry", guess_max=20000)

 auc23 <- readxl::read_excel("//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/WCVI/ESCAPEMENT/Data/2023/SILs/Area 21-22/Nitinat River/A22_Nitinat_River_2023.xlsx", sheet="AUC_R", guess_max=20000)

tribs.sil <- readxl::read_excel("//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/WCVI/ESCAPEMENT/Data/2023/SILs/Area 21-22/Tribs/A22_Tributaries_2023.xlsx", sheet="Tributaries Raw SIL entry", guess_max=20000)


# SIL+AUC database ----------------------------
ver23 <- readxl::read_excel(path=paste0(OD.dir, "SIL+AUC_database_output_2024-01-03 - AUC.xlsx"), sheet="raw")
AUC23 <- readxl::read_excel(path=paste0(OD.dir, "SIL+AUC_database_output_2024-01-03 - AUC.xlsx"), sheet="AUC")


# Hatchery ----------------------------
nrh.cn <- readxl::read_excel("//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/WCVI/ESCAPEMENT/Data/2023/SILs/Area 21-22/Adult_Management - Nitinat CN 2023.xlsx", sheet="Adult Management", guess_max=20000, skip=1) %>% 
  rename(InventoryLot=`Inventory Lot`, 
         ParentActivityType=`Parent Activity Type`, 
         SourceLocationType=`Source Location Type`,
         TotalCount=`Total Count`,
         MaturityClass=`Maturity Class`)

nrh.cm <- readxl::read_excel("//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/WCVI/ESCAPEMENT/Data/2023/SILs/Area 21-22/Nitinat Chum Removals_By Date_Nov 29_2023.xlsx", sheet="All Adult Activities", skip=1) 


# Enviro ----------------------------
nit.water <- list.files(here::here("data", "enviro"), pattern = "NitinatWaterLevelHydrometData", full.names = T) %>% 
  map(~readxl::read_excel(path=.x, sheet="Exported Data", guess_max=20000)) %>%
  purrr::list_rbind() %>%
  mutate(gauge_m = case_when(grepl("Estimated", `Staff Gauge(mH20)`) ~ NA,
                             TRUE ~ as.numeric(`Staff Gauge(mH20)`)),
         sensor_depth = case_when(grepl("Estimated", `Sensor Depth(mH20)`) ~ NA,
                             TRUE ~ as.numeric(`Sensor Depth(mH20)`)),
         date = as.Date(Time)) %>% 
  filter(date<= as.Date("2023-11-08"), gauge_m>0.000)



# Run timing ----------------------------
nit.rtStWk <- readxl::read_excel("//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/WCVI/ESCAPEMENT/Data/run timing tables (updated).xlsx", 
                                    sheet="Nitinat", guess_max=10000, trim_ws=T, range=cell_limits(c(4, 1), c(NA, NA))) %>% 
  select(-c(`...24`:`...75`)) %>%
  rename(year=`...1`) %>% 
  filter(if_any(everything(), ~ !is.na(.)), !grepl("Table", year), !grepl("Year", year), !is.na(year)) %>% 
  mutate(species = c(rep("chinook", nrow(.)/3), rep("coho", nrow(.)/3), rep("chum", nrow(.)/3)),
         `92` = case_when(`92`=="1000-2000" ~ as.numeric(1500),
                          TRUE ~ as.numeric(`92`)),
         across(c(`81`:`124`), ~as.numeric(.)),
         plot_group = case_when(year==analysis_year ~ year,
                                year!=analysis_year ~ paste0(min(year), sep="-", (analysis_year-1)))) %>%
  pivot_longer(cols = c(`81`:`124`), names_to = "stat_week", values_to = "count") %>% 
  print()

nit.rtStWk$stat_week <- factor(nit.rtStWk$stat_week,
                               levels=c("81","82","83","84",
                                        "91","92","93","94",
                                        "101","102","103","104","105",
                                        "111","112","113","114","115",
                                        "121","122","123","124", ordered=T))

# Escapement ----------------------------
nit.esc <- readxl::read_excel(path=here("data", "escapement", "NuSEDS_Area21-22_escapement_Jan2024.xlsx"), sheet="Data")

nit.esc.goal <- read.csv(here("data", "escapement", "wcviCK-BootstrappedRPs.csv")) %>% 
  filter(Stock=="Nitinat")
```

<br>

<br>


# **Chinook**

Six snorkel surveys and one helicopter survey were conducted for Chinook in 2023. 

<br>

## Raw vs expanded counts

Counts are expanded for both % habitat seen and observer efficiency, depending on survey conditions. 

```{r}
ggplot() +
  geom_bar(data=ver23 %>% filter(ShortSpecies=="CN") %>% pivot_longer(LiveAdults:ExpDeadJacks, names_to="est_type", values_to="n") %>% filter(est_type %in% c("LiveAdults", "ExpLiveAdults")), 
           aes(x=as.Date(Date), y=n, colour=est_type, fill=est_type), stat="identity", position="dodge", alpha=0.5) +
  labs(x="", y="Adult Chinook counts", fill="Expansion", colour="Expansion") +
  scale_x_date(date_breaks="7 day", date_labels = "%b %d") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

Figure: Live Adult Chinook counts already expanded for % habitat seen (blue), and with additional % population expansion (pink). 

<br>

## Counts & environmentals

```{r }
ggplot() +
  geom_ribbon(data=nit.water %>% group_by(date) %>% summarize(min=min(gauge_m), max=max(gauge_m)), 
              aes(x=as.Date(date), ymin=min*5000, ymax=max*5000), fill="dodger blue", alpha=0.5) +
  geom_line(data=nit.water %>% group_by(date) %>% summarize(mean = mean(gauge_m)), 
            aes(x=as.Date(date), y=mean*5000), colour="dodger blue") +
  geom_bar(data=nit.sil %>% filter(species=="Chinook") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
             summarize(n=sum(`adults live / habitat seen (# Est Live)`)), 
           aes(x=as.Date(date), y=n, group=type, fill=`upper/lower (for Nitinat River only)`, colour=`upper/lower (for Nitinat River only)`), 
           stat="identity", position="stack", alpha=0.7) +
  # geom_line(data=nit.sil %>% filter(species=="Chinook") %>% group_by(date) %>% 
  #             summarize(n=sum(`adults dead`)), 
  #           aes(x=as.Date(date), y=n), 
  #           stat="identity", position="stack") +
  labs(x="", y="Expanded adult Chinook") +
  scale_y_continuous(sec.axis = sec_axis(~./5000, name="Staff gauge (m)")) +
  scale_x_date(date_breaks="7 day", date_labels="%b %d") +
  labs(x="", Y="Live chinook", fill="Segment", colour="Segment") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

Figure: Chinook counts (expanded adults) and water levels (blue line) 

<br>

<br>

## Behaviour & carcasses

```{r}
  # ggplot() +
  #   geom_bar(data=nit.sil %>% filter(species=="Chinook") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
  #              summarize(n=sum(`adults live / habitat seen (# Est Live)`)), 
  #            aes(x=as.Date(date), y=n, fill=`upper/lower (for Nitinat River only)`, colour=`upper/lower (for Nitinat River only)`), 
  #            stat="identity", position="stack", alpha=0.7) +
  #   geom_point(data=nit.sil %>% filter(species=="Chinook") %>% group_by(date) %>% 
  #                summarize(n=sum(`adults dead`)), 
  #              aes(x=as.Date(date), y=n*10), size=2, alpha=0.5) +
  #   geom_line(data=nit.sil %>% filter(species=="Chinook") %>% group_by(date) %>% 
  #               summarize(n=sum(`adults dead`)), 
  #             aes(x=as.Date(date), y=n*10), 
  #             stat="identity", position="stack") +
  #   scale_y_continuous(sec.axis = sec_axis(~./10, name="Carcasses")) +
  #   scale_x_date(date_breaks="7 day", date_labels = "%b %d")+
  #   labs(x="", y="Expanded live adult Chinook", fill="Segment", colour="Segment") +
  #   theme_bw() +
  #   theme(axis.text.x = element_text(angle=45, hjust=1)),
  
  ggplot() +
    geom_bar(data=nit.sil %>% filter(species=="Chinook") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
               summarize(Spawning=sum(`adults # spawning`),
                         Holding = sum(`adults # holding`)) %>% pivot_longer(cols=c(Spawning,Holding), names_to="behaviour", values_to="count"), 
             aes(x=as.Date(date), y=count, fill=behaviour, colour=`upper/lower (for Nitinat River only)`), 
             stat="identity", position="stack", alpha=0.7, size=1) +
    geom_point(data=nit.sil %>% filter(species=="Chinook") %>% group_by(date) %>% 
                 summarize(n=sum(`adults dead`)), 
               aes(x=as.Date(date), y=n*10), size=2, alpha=0.5) +
    geom_line(data=nit.sil %>% filter(species=="Chinook") %>% group_by(date) %>% 
                summarize(n=sum(`adults dead`)), 
              aes(x=as.Date(date), y=n*10)) +
    scale_y_continuous(breaks=seq(0,20000,by=2500), sec.axis = sec_axis(~./10, name="Carcasses")) +
    scale_x_date(date_breaks="7 day", date_labels = "%b %d")+
    scale_fill_manual(breaks=waiver(), values=c("transparent", "gray")) +
    labs(x="", y="Spawning adult Chinook", fill="Behaviour", colour="Segment") +
    theme_bw() +
    theme(axis.text.x = element_text(angle=45, hjust=1))
```
<br>

Figure: Expanded live adult Chinook (bars) and carcasses (black line). Bar fill indicates behaviour (spawning=gray, holding=white), bar outline indicates river segment (upper=blue, lower=pink).

<br>

<br>

## Snorkel & swim-in counts

```{r}
  # ggplot() +
  #   geom_bar(data=nit.sil %>% filter(species=="Chinook") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
  #              summarize(n=sum(`adults live / habitat seen (# Est Live)`)), 
  #            aes(x=as.Date(date), y=n, group=type, fill=`upper/lower (for Nitinat River only)`, colour=`upper/lower (for Nitinat River only)`), 
  #            stat="identity", position="stack", alpha=0.7) + 
  #   geom_bar(data=nrh.cn %>% filter(`Parent Activity Type`=="Adult Capture", grepl("Swim Ins", `Inventory Lot`)) %>% 
  #              group_by(`Activity Start Date`) %>% summarize(n=sum(`Total Count`)),
  #            aes(x=as.Date(`Activity Start Date`), y=n*1), fill="black", stat="identity", alpha=0.7) +
  #   scale_y_continuous(sec.axis = sec_axis(~./1, name="Swim ins")) +
  #   scale_x_date(date_breaks="7 day", date_labels = "%b %d")+
  #   labs(x="", y="Expanded adult Chinook", fill="Segment", colour="Segment") +
  #   theme_bw() +
  #   theme(axis.text.x = element_text(angle=45, hjust=1)),
  
  ggplot() +
    geom_bar(data=nit.sil %>% filter(species=="Chinook") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
               summarize(Spawning=sum(`adults # spawning`),
                         Holding = sum(`adults # holding`)) %>% pivot_longer(cols=c(Spawning,Holding), names_to="behaviour", values_to="count"), 
             aes(x=as.Date(date), y=count, fill=behaviour, colour=`upper/lower (for Nitinat River only)`), 
             stat="identity", position="stack", alpha=0.7, size=1) +
    geom_bar(data=nrh.cn %>% filter(ParentActivityType=="Adult Capture", grepl("Swim Ins", InventoryLot)) %>% 
               group_by(`Activity Start Date`) %>% summarize(n=sum(TotalCount)),
             aes(x=as.Date(`Activity Start Date`), y=n*1), fill="black", stat="identity", alpha=0.7) +
    scale_y_continuous(sec.axis = sec_axis(~./1, name="Swim ins")) +
    scale_x_date(date_breaks="7 day", date_labels = "%b %d") +
    scale_fill_manual(breaks=waiver(), values=c("transparent", "gray")) +
    labs(x="", y="Spawning adult Chinook", fill="Behaviour", colour="Behaviour") +
    theme_bw() +
    theme(axis.text.x = element_text(angle=45, hjust=1))
```

Figure: Expanded adult Chinook counts (large bars) and swim-ins (narrow black bars). Adult bar fill colours indicate behaviour (spawning=gray, holding=white), bar line colour indicates river segment (lower=pink, upper=blue).

<br>

<br>

## Preliminary escapement estimate (AUC/PL+D) 

<br>

Table: Raw chinook counts

```{r}
ver23 %>% 
  filter(ShortSpecies=="CN") %>% 
  select(-c(FabricatedZero)) %>% 
  mutate(across(c(LiveAdults:ExpDeadJacks), ~round(., 0))) %>% 
  kbl(align="c") %>%
  kable_paper("hover", full_width=T, position = "center") %>%
  collapse_rows(columns=c(1:4), valign="middle", target=1) 
```

<br>

```{r}
ggarrange(
ggplot() +
  geom_line(data=AUC23%>%filter(estMethod=="with heli", species=="Chinook"), aes(x=as.Date(Date), y=ExpAdultFishDays)) +
  geom_point(data=AUC23%>%filter(estMethod=="with heli", species=="Chinook"), 
             aes(x=as.Date(Date), y=ExpAdultFishDays, fill=surveyMethod, colour=surveyMethod), shape=21, size=4) +
  scale_fill_manual(breaks=c("heli", "snorkel"), values=c("orange", "blue")) +
  scale_colour_manual(breaks=c("heli", "snorkel"), values=c("orange", "blue")) +
  labs(x="", y="Fish-days", fill="", colour="") +
  theme_bw(),

ggplot() +
  geom_line(data=AUC23%>%filter(estMethod=="without heli", species=="Chinook"), aes(x=as.Date(Date), y=ExpAdultFishDays)) +
  geom_point(data=AUC23%>%filter(estMethod=="without heli", species=="Chinook"), 
             aes(x=as.Date(Date), y=ExpAdultFishDays, fill=surveyMethod, colour=surveyMethod), shape=21, size=4) +
  scale_fill_manual(breaks=c("heli", "snorkel"), values=c("orange", "blue")) +
  scale_colour_manual(breaks=c("heli", "snorkel"), values=c("orange", "blue")) +
  labs(x="", y="Fish-days", fill="", colour="") +
  theme_bw(),

nrow=2,
common.legend = T, legend = "bottom"
)
```

Figure: Chinook AUC curves with heli (top) and without heli (bottom) survey. 

<br>

Table: AUC math (SIL+AUC database math extracted: SIL+AUC_database_output_2024-01-03 - AUC.xlsx)

```{r}
AUC23 %>% 
  filter(species=="Chinook") %>% 
  relocate(surveyMethod, .before=Date) %>% 
  kbl(align="c") %>%
  kable_paper("hover", full_width=T, position = "center") %>%
  collapse_rows(columns=c(1:4), valign="middle", target=1) 
```

<br>

Table: Escapement estimates with/without helicopter survey for each survey life value.
```{r}
AUC23 %>% 
  filter(species=="Chinook") %>%
  select(-c(species)) %>%
  group_by(estMethod) %>% 
  summarize(fishDays = sum(ExpAdultFishDays),
            est_SL15 = round(fishDays/15,0),
            est_SL20 = round(fishDays/20,0),
            est_SL25 = round(fishDays/25,0)) %>% 
  pivot_longer(cols=c(est_SL15:est_SL25), names_to = "Rank", values_to="Estimate") %>% 
  mutate(Rank = case_when(grepl("15", Rank) ~ "Low Value",
                           grepl("20", Rank) ~ "Middle Value",
                           grepl("25", Rank) ~ "High Value"),
         SL = case_when(grepl("Low", Rank) ~ 15,
                        grepl("Middle", Rank) ~ 20,
                        grepl("High", Rank) ~ 25),
         Est_without_swimins = Estimate - (nrh.cn %>% 
              filter(grepl("Swim In", InventoryLot) & ParentActivityType=="Adult Capture" & SourceLocationType=="Brailer" & 
                       MaturityClass%in%c("Male","Female")) %>% 
              summarize(n=sum(TotalCount)) %>%
              pull(n))) %>%
  relocate(SL, .after = Rank) %>%
  kbl(align="c") %>%
  kable_paper("hover", full_width=T, position = "center") %>%
  collapse_rows(columns=c(1:4), valign="middle", target=1) 
```

<br>

### Estimate decision:   

For 2023 estimate we went with expand AUC SL=15 for three reasons: First, there were a lot more swim-ins than usual this year, which was attributed to the COVID releases leaving the hatchery directly rather than being reared in the lake pen. Fish homing to the hatchery may have had more directed migration and therefore not stayed in the system long. Second, the higher estimate from the shorter SL resulted in an estimate that was high enough to cover the raw peak live count of ~15k,  but once removing the `r nrh.cn %>% filter(grepl("Swim In", InventoryLot) & ParentActivityType=="Adult Capture" & SourceLocationType=="Brailer" & MaturityClass %in% c("Male","Female")) %>% summarize(n=sum(TotalCount)) %>% pull(n)` adult and `r nrh.cn %>% filter(grepl("Swim In", InventoryLot) & ParentActivityType=="Adult Capture" & SourceLocationType=="Brailer" & MaturityClass=="Jack") %>% summarize(n=sum(TotalCount)) %>% pull(n)` jack swim-ins, still resulted in a reasonable estimate that agreed with what the hatchery observed. Third, there may have been a slight delay in entry as the water was low for the early season and rains were a bit later than usual (though not as late as 2022). The Expanded AUC adult estimate for 2023 spawners = ~21098 before swim-ins, ~14252 after swimins removed; expanded jacks = ~107 before swim-ins, removal of swim-ins results in 0 jacks.  *This is subject to change; see EscBatchHardCopy report for specific estimate abundance.*  

<br>

<br>

## Tributaries

Chinook were not seen in any Nitinat River or Lake tributaries in 2023.

```{r}
tribs.sil %>% 
  filter(species=="Chinook") %>% 
  group_by(system) %>% 
  summarize(exp_ad_live = sum(`adults live / habitat seen (# Est Live)`)) %>%
  kbl(align="c") %>%
  kable_paper("hover", full_width=T, position = "center") 
```

<br>

<br>

## `r analysis_year` compared to historical run timing and abundances

```{r}
ggplot() +
  geom_point(data=nit.rtStWk%>%filter(species=="chinook"),
           aes(x=stat_week, y=count, colour=plot_group, fill=plot_group, size=plot_group),
           shape=21, alpha=0.7) +
  scale_fill_manual(breaks=waiver(), values=c("gray60", "dodger blue")) +
  scale_colour_manual(breaks=waiver(), values=c("gray60", "dodger blue")) +
  scale_size_manual(breaks=waiver(), values=c(1.5,3.5)) +
  labs(x="Stat week", y="Live adult Chinook", fill="Year:", colour="Year:", size="Year:") +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.box = "horizontal")
```

Figure: Return timing of Chinook in 2023 relative to recent returns. 

<br>

```{r}
ggplot()  +
  geom_hline(data=nit.esc.goal %>% filter(RP=="SMSY"), aes(yintercept=Value)) +
  geom_bar(data=nit.esc %>% filter(`Waterbody Name`=="NITINAT RIVER", `Species Qualified`=="CK"), 
           aes(x=`Analysis Year`, y=`Natural Adult Spawners`), stat="identity", colour="dodger blue", fill="dodger blue", alpha=0.8) +
  geom_bar(data=nit.esc %>% filter(`Waterbody Name`=="NITINAT RIVER", `Species Qualified`=="CK"), 
           aes(x=`Analysis Year`, y=`Unspecified Return`), stat="identity", colour="gray80", fill="gray80", alpha=0.8) +  
  geom_bar(aes(x=analysis_year, y=14252), stat="identity", colour="orange", fill="orange", alpha=0.8) +   
  scale_x_continuous(breaks=seq(min(nit.esc$`Analysis Year`), analysis_year, by=2)) +
  labs(x="Return year", y="Chinook spawners") +
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

Figure: 2023 preliminary adult Chinook escapement estimate (orange) relative to historical adult (blue) and unspecified (gray) escapement estimates.

<br>

<br>

<br>













# **Chum** 

<br>

Six snorkel surveys and one helicopter survey were conducted for Chum in 2023. The heli survey may have been slightly early for Chum. 

<br>

## Raw vs expanded counts

Counts are expanded for both % habitat seen and observer efficiency, depending on survey conditions. 

```{r}
ggplot() +
  geom_bar(data=ver23 %>% filter(ShortSpecies=="CM") %>% pivot_longer(LiveAdults:ExpDeadJacks, names_to="est_type", values_to="n") %>% filter(est_type %in% c("LiveAdults", "ExpLiveAdults")), 
           aes(x=as.Date(Date), y=n, colour=est_type, fill=est_type), stat="identity", position="dodge", alpha=0.5) +
  labs(x="", y="Adult Chum counts", fill="Expansion", colour="Expansion") +
  scale_x_date(date_breaks="7 day", date_labels = "%b %d") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

Figure: Live Adult Chum counts already expanded for % habitat seen (blue), and with additional % population expansion (pink). 

<br>

## Counts & environmentals

```{r}
ggplot() +
  geom_ribbon(data=nit.water %>% group_by(date) %>% summarize(min=min(gauge_m), max=max(gauge_m)), 
              aes(x=as.Date(date), ymin=min*5000, ymax=max*5000), fill="dodger blue", alpha=0.5) +
  geom_line(data=nit.water %>% group_by(date) %>% summarize(mean = mean(gauge_m)), 
            aes(x=as.Date(date), y=mean*5000), colour="dodger blue") +
  geom_bar(data=nit.sil %>% filter(species=="Chum") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
             summarize(n=sum(`adults live / habitat seen (# Est Live)`)), 
           aes(x=as.Date(date), y=n, group=type, fill=`upper/lower (for Nitinat River only)`, colour=`upper/lower (for Nitinat River only)`), 
           stat="identity", position="stack", alpha=0.7) +
  # geom_line(data=nit.sil %>% filter(species=="Chum") %>% group_by(date) %>% 
  #             summarize(n=sum(`adults dead`)), 
  #           aes(x=as.Date(date), y=n), 
  #           stat="identity", position="stack") +
  scale_y_continuous(sec.axis = sec_axis(~./5000, name="Staff gauge (m)")) +
  scale_x_date(date_breaks="7 day", date_labels="%b %d") +
  labs(x="", y="Expanded adult Chum", fill="Segment", colour="Segment") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

Figure: Chinook counts (expanded adults) and water levels (blue line) 

<br>

<br>

## Behaviour & carcasses

```{r}
  # ggplot() +
  #   geom_bar(data=nit.sil %>% filter(species=="Chum") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
  #              summarize(n=sum(`adults live / habitat seen (# Est Live)`)), 
  #            aes(x=as.Date(date), y=n, fill=`upper/lower (for Nitinat River only)`, colour=`upper/lower (for Nitinat River only)`), 
  #            stat="identity", position="stack", alpha=0.7) +
  #   geom_point(data=nit.sil %>% filter(species=="Chum") %>% group_by(date) %>% 
  #                summarize(n=sum(`adults dead`)), 
  #              aes(x=as.Date(date), y=n*10), size=2, alpha=0.5) +
  #   geom_line(data=nit.sil %>% filter(species=="Chum") %>% group_by(date) %>% 
  #               summarize(n=sum(`adults dead`)), 
  #             aes(x=as.Date(date), y=n*10), 
  #             stat="identity", position="stack") +
  #   scale_y_continuous(sec.axis = sec_axis(~./10, name="Carcasses")) +
  #   scale_x_date(date_breaks="7 day", date_labels = "%b %d")+
  #   labs(x="", y="Expanded live adult Chum", fill="Segment", colour="Segment") +
  #   theme_bw() +
  #   theme(axis.text.x = element_text(angle=45, hjust=1)),
  
  ggplot() +
    geom_bar(data=nit.sil %>% filter(species=="Chum") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
               summarize(Spawning=sum(`adults # spawning`),
                         Holding = sum(`adults # holding`)) %>% pivot_longer(cols=c(Spawning,Holding), names_to="behaviour", values_to="count"), 
             aes(x=as.Date(date), y=count, fill=behaviour, colour=`upper/lower (for Nitinat River only)`), 
             stat="identity", position="stack", alpha=0.7, size=1) +
    geom_point(data=nit.sil %>% filter(species=="Chum") %>% group_by(date) %>% 
                 summarize(n=sum(`adults dead`)), 
               aes(x=as.Date(date), y=n*10), size=2, alpha=0.5) +
    geom_line(data=nit.sil %>% filter(species=="Chum") %>% group_by(date) %>% 
                summarize(n=sum(`adults dead`)), 
              aes(x=as.Date(date), y=n*10), 
              stat="identity", position="stack") +
    scale_y_continuous(breaks=seq(0,20000, by=2500), sec.axis = sec_axis(~./10, name="Carcasses")) +
    scale_x_date(date_breaks="7 day", date_labels = "%b %d")+
    scale_fill_manual(breaks=waiver(), values=c("transparent", "gray")) +
    labs(x="", y="Spawning adult Chum", fill="Behaviour", colour="Segment") +
    theme_bw() +
    theme(axis.text.x = element_text(angle=45, hjust=1))
```
<br>

Figure: Expanded live adult Chum (bars) and carcasses (black line). Bar fill indicates behaviour (spawning=gray, holding=white), bar outline indicates river segment (upper=blue, lower=pink)

<br>

<br>

## Snorkel & swim-in counts

```{r}
  # ggplot() +
  #   geom_bar(data=nit.sil %>% filter(species=="Chum") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
  #              summarize(n=sum(`adults live / habitat seen (# Est Live)`)), 
  #            aes(x=as.Date(date), y=n, group=type, fill=`upper/lower (for Nitinat River only)`, colour=`upper/lower (for Nitinat River only)`), 
  #            stat="identity", position="stack", alpha=0.7) + 
  #   geom_bar(data=nrh.cm, 
  #            aes(x=as.Date(`Start Date`), y=`Total Count`*5), fill="black", stat="identity", alpha=0.7) +
  #   scale_y_continuous(sec.axis = sec_axis(~./5, name="Swim ins")) +
  #   scale_x_date(date_breaks="7 day", date_labels = "%b %d")+
  #   labs(x="", y="Expanded adult Chum", fill="Segment", colour="Segment") +
  #   theme_bw() +
  #   theme(axis.text.x = element_text(angle=45, hjust=1)),
  
  ggplot() +
    geom_bar(data=nit.sil %>% filter(species=="Chum") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
               summarize(Spawning=sum(`adults # spawning`),
                         Holding = sum(`adults # holding`)) %>% pivot_longer(cols=c(Spawning,Holding), names_to="behaviour", values_to="count"), 
             aes(x=as.Date(date), y=count, fill=behaviour, colour=`upper/lower (for Nitinat River only)`), 
             stat="identity", position="stack", alpha=0.7, size=1) +
    geom_bar(data=nrh.cm,
             aes(x=as.Date(`Start Date`), y=`Total Count`*5), fill="black", stat="identity", alpha=0.7) +
    scale_y_continuous(sec.axis = sec_axis(~./5, name="Swim ins")) +
    scale_x_date(date_breaks="7 day", date_labels = "%b %d") +
    scale_fill_manual(breaks=waiver(), values=c("transparent", "gray")) +
    labs(x="", y="Spawning adult Chum", fill="Behaviour", colour="Behaviour") +
    theme_bw() +
    theme(axis.text.x = element_text(angle=45, hjust=1))
```

Figure: Expanded adult Chum counts (large bars) and swim-ins (narrow black bars). Adult bar fill colours indicate behaviour (spawning=gray, holding=white), bar line colour indicates river segment (lower=pink, upper=blue).

<br>

<br>

## Preliminary escapement estimate (AUC) 

<br>

Table: Raw chum counts

```{r}
ver23 %>% 
  filter(ShortSpecies=="CM") %>% 
  select(-c(FabricatedZero)) %>% 
  mutate(across(c(LiveAdults:ExpDeadJacks), ~round(., 0))) %>% 
  kbl(align="c") %>%
  kable_paper("hover", full_width=T, position = "center") %>%
  collapse_rows(columns=c(1:4), valign="middle", target=1)
```

<br>

```{r}
ggarrange(
ggplot() +
  geom_line(data=AUC23%>%filter(estMethod=="with heli", species=="Chum"), aes(x=as.Date(Date), y=ExpAdultFishDays)) +
  geom_point(data=AUC23%>%filter(estMethod=="with heli", species=="Chum"), 
             aes(x=as.Date(Date), y=ExpAdultFishDays, fill=surveyMethod, colour=surveyMethod), shape=21, size=4) +
  scale_fill_manual(breaks=c("heli", "snorkel"), values=c("orange", "blue")) +
  scale_colour_manual(breaks=c("heli", "snorkel"), values=c("orange", "blue")) +
  labs(x="", y="Fish-days", fill="", colour="") +
  theme_bw(),

ggplot() +
  geom_line(data=AUC23%>%filter(estMethod=="without heli", species=="Chum"), aes(x=as.Date(Date), y=ExpAdultFishDays)) +
  geom_point(data=AUC23%>%filter(estMethod=="without heli", species=="Chum"), 
             aes(x=as.Date(Date), y=ExpAdultFishDays, fill=surveyMethod, colour=surveyMethod), shape=21, size=4) +
  scale_fill_manual(breaks=c("heli", "snorkel"), values=c("orange", "blue")) +
  scale_colour_manual(breaks=c("heli", "snorkel"), values=c("orange", "blue")) +
  labs(x="", y="Fish-days", fill="", colour="") +
  theme_bw(),
nrow=2,
common.legend = T, legend = "bottom"
)
```

Figure: AUC curves with heli (top) and without heli (bottom) survey. 

<br>

Table: AUC math (SIL+AUC database math extracted: SIL+AUC_database_output_2024-01-03 - AUC.xlsx)

```{r}
AUC23 %>% 
  filter(species=="Chum") %>% 
  relocate(surveyMethod, .before=Date) %>% 
  kbl(align="c") %>%
  kable_paper("hover", full_width=T, position = "center") %>%
  collapse_rows(columns=c(1:4), valign="middle", target=1) 
```

<br>

Table: Escapement estimates with/without helicopter survey for each survey life value.

```{r}
# auc23 %>% 
#   filter(species=="chum") %>%
#   group_by(method) %>% 
#   summarize(fish_days_SL8 = round(sum(`expanded adult`)/8,0),
#             fish_days_SL10 = round(sum(`expanded adult`)/10,0),
#             fish_days_SL15 = round(sum(`expanded adult`)/15,0)) %>% 
#   pivot_longer(cols=c(fish_days_SL8:fish_days_SL15), names_to = "Rank", values_to="Estimate") %>% 
#   mutate(Rank = case_when(grepl("8", Rank) ~ "Low Value",
#                            grepl("10", Rank) ~ "Middle Value",
#                            grepl("15", Rank) ~ "High Value"),
#          SL = case_when(grepl("Low", Rank) ~ 8,
#                         grepl("Middle", Rank) ~ 10,
#                         grepl("High", Rank) ~ 15),
#          Est_without_swimins = 
#            Estimate - 
#            nrh.cm %>% 
#               summarize(n=sum(TotalCount)) %>%
#               pull(n)
#            ) %>%
#   relocate(SL, .after = Rank) %>%
#   kbl(align="c") %>%
#   kable_paper("hover", full_width=T, position = "center") %>%
#   collapse_rows(columns=c(1:4), valign="middle", target=1) 

AUC23 %>% 
  filter(species=="Chum") %>%
  select(-c(species)) %>%
  group_by(estMethod) %>% 
  summarize(fishDays = sum(ExpAdultFishDays),
            est_SLL = round(fishDays/10,0),
            est_SLM = round(fishDays/15,0),
            est_SLH = round(fishDays/20,0)) %>% 
  pivot_longer(cols=c(est_SLL:est_SLH), names_to = "Rank", values_to="Estimate") %>% 
  mutate(Rank = case_when(grepl("SLL", Rank) ~ "Low Value",
                           grepl("SLM", Rank) ~ "Middle Value",
                           grepl("SLH", Rank) ~ "High Value"),
         SL = case_when(grepl("Low", Rank) ~ 10,
                        grepl("Middle", Rank) ~ 15,
                        grepl("High", Rank) ~ 20)) %>%
  relocate(SL, .after = Rank) %>%
  kbl(align="c") %>%
  kable_paper("hover", full_width=T, position = "center") %>%
  collapse_rows(columns=c(1:4), valign="middle", target=1) 
```

<br>

### Estimate decision:

For `r analysis_year` Chum estimate, needs review with NRH. Shorter SL is very expanded. PL+D gives 15222 for reference. 

<br>

<br>

## Tributaries

Live chum were present in Campus and Hobiton Creek in 2023, and dead chum were noted in NoName Creek. Opportunistic surveys of Caycuse and Doobah found no chum. 

```{r}
tribs.sil %>% 
  filter(species=="Chum") %>% 
  group_by(system) %>% 
  summarize(exp_ad_live = sum(`adults live / habitat seen (# Est Live)`),
            dead_adults = sum(`adults dead`)) %>%
  mutate(across(c(exp_ad_live:dead_adults), ~round(., 0))) %>%
  kbl(align="c") %>%
  kable_paper("hover", full_width=T, position = "center") %>%
  collapse_rows(columns=c(1), valign="middle", target=1) 
```

<br>

<br>

## `r analysis_year` compared to historical run timing

```{r}
ggplot() +
  geom_point(data=nit.rtStWk%>%filter(species=="chum"),
           aes(x=stat_week, y=count, colour=plot_group, fill=plot_group, size=plot_group),
           shape=21, alpha=0.7) +
  scale_fill_manual(breaks=waiver(), values=c("gray60", "dodger blue")) +
  scale_colour_manual(breaks=waiver(), values=c("gray60", "dodger blue")) +
  scale_size_manual(breaks=waiver(), values=c(1.5,3.5)) +
  labs(x="Stat week", y="Live adult Coho", fill="Year:", colour="Year:", size="Year:") +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.box = "horizontal")
```

<br>

```{r}
ggplot()  +
  geom_bar(data=nit.esc %>% filter(`Waterbody Name`=="NITINAT RIVER", `Species Qualified`=="CM"), 
           aes(x=`Analysis Year`, y=`Natural Adult Spawners`), stat="identity", colour="dodger blue", fill="dodger blue", alpha=0.8) +
  geom_bar(data=nit.esc %>% filter(`Waterbody Name`=="NITINAT RIVER", `Species Qualified`=="CM"), 
           aes(x=`Analysis Year`, y=`Unspecified Return`), stat="identity", colour="gray80", fill="gray80", alpha=0.8) +  
  geom_bar(aes(x=analysis_year, y=30000), stat="identity", colour="orange", fill="orange", alpha=0.8) +   
  scale_x_continuous(breaks=seq(min(nit.esc$`Analysis Year`), analysis_year, by=2)) +
  labs(x="Return Year", y="Chum spawners") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

Figure: 2023 preliminary adult Chum escapement estimate (orange) relative to historical adult (blue) and unspecified (gray) escapement estimates. There was no Chum estimate available in 1990. 

<br>

<br>

<br>


















# **Coho**            
 
<br>

Six snorkel surveys were conducted in 2023, but only the latter ones can be considered representative for Coho (they were only seen on 4 of the 6 surveys).  

<br>

## Raw vs expanded counts

Counts are expanded for both % habitat seen and observer efficiency, depending on survey conditions. 

```{r}
ggplot() +
  geom_bar(data=ver23 %>% filter(ShortSpecies=="CO") %>% pivot_longer(LiveAdults:ExpDeadJacks, names_to="est_type", values_to="n") %>% filter(est_type %in% c("LiveAdults", "ExpLiveAdults")), 
           aes(x=as.Date(Date), y=n, colour=est_type, fill=est_type), stat="identity", position="dodge", alpha=0.5) +
  labs(x="", y="Adult Coho counts", fill="Expansion", colour="Expansion") +
  scale_x_date(date_breaks="7 day", date_labels = "%b %d") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

Figure: Live Adult Coho counts already expanded for % habitat seen (blue), and with additional % population expansion (pink). 

<br>

## Counts & environmentals

```{r eval=F,include=F}
ggplot() +
  geom_ribbon(data=nit.water %>% group_by(date) %>% summarize(min=min(gauge_m), max=max(gauge_m)), 
              aes(x=as.Date(date), ymin=min*100, ymax=max*100), fill="dodger blue", alpha=0.5) +
  geom_line(data=nit.water %>% group_by(date) %>% summarize(mean = mean(gauge_m)), 
            aes(x=as.Date(date), y=mean*100), colour="dodger blue") +
  geom_bar(data=nit.sil %>% filter(species=="Coho") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
             summarize(n=sum(`adults live / habitat seen (# Est Live)`)), 
           aes(x=as.Date(date), y=n, group=type, fill=`upper/lower (for Nitinat River only)`, colour=`upper/lower (for Nitinat River only)`), 
           stat="identity", position="stack", alpha=0.7) +
  # geom_line(data=nit.sil %>% filter(species=="Coho") %>% group_by(date) %>% 
  #             summarize(n=sum(`adults dead`)), 
  #           aes(x=as.Date(date), y=n), 
  #           stat="identity", position="stack") +
  scale_y_continuous(sec.axis = sec_axis(~./100, name="Staff gauge (m)")) +
  scale_x_date(date_breaks="7 day", date_labels="%b %d") +
  labs(x="", y="Expanded adult Coho", fill="Segment", colour="Segment") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

Figure: Coho counts (expanded adults) and water levels (blue line) 

<br>

<br>

## Behaviour & carcasses

```{r eval=F,include=F}
ggplot() +
  geom_bar(data=nit.sil %>% filter(species=="Coho") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
             summarize(Spawning=sum(`adults # spawning`),
                       Holding = sum(`adults # holding`)) %>% pivot_longer(cols=c(Spawning,Holding), names_to="behaviour", values_to="count"), 
           aes(x=as.Date(date), y=count, fill=behaviour, colour=`upper/lower (for Nitinat River only)`), 
           stat="identity", position="stack", alpha=0.7, size=1) +
  geom_point(data=nit.sil %>% filter(species=="Coho") %>% group_by(date) %>% 
               summarize(n=sum(`adults dead`)), 
             aes(x=as.Date(date), y=n*10), size=2, alpha=0.5) +
  geom_line(data=nit.sil %>% filter(species=="Coho") %>% group_by(date) %>% 
              summarize(n=sum(`adults dead`)), 
            aes(x=as.Date(date), y=n*10), 
            stat="identity", position="stack") +
  scale_y_continuous(sec.axis = sec_axis(~./10, name="Carcasses")) +
  scale_x_date(date_breaks="7 day", date_labels = "%b %d")+
  scale_fill_manual(breaks=waiver(), values=c("transparent", "gray")) +
  labs(x="", y="Spawning adult Coho", fill="Behaviour", colour="Segment") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
```
<br>

Figure: Expanded live adult Coho (bars) and carcasses (black line). Bar fill indicates behaviour (spawning=gray, holding=white), bar outline indicates river segment (upper=blue, lower=pink).

<br>

<br>

## Snorkel & swim-in counts

```{r eval=F,include=F}
  # ggplot() +
  #   geom_bar(data=nit.sil %>% filter(species=="Chum") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
  #              summarize(n=sum(`adults live / habitat seen (# Est Live)`)), 
  #            aes(x=as.Date(date), y=n, group=type, fill=`upper/lower (for Nitinat River only)`, colour=`upper/lower (for Nitinat River only)`), 
  #            stat="identity", position="stack", alpha=0.7) + 
  #   geom_bar(data=nrh.cm, 
  #            aes(x=as.Date(`Start Date`), y=`Total Count`*5), fill="black", stat="identity", alpha=0.7) +
  #   scale_y_continuous(sec.axis = sec_axis(~./5, name="Swim ins")) +
  #   scale_x_date(date_breaks="7 day", date_labels = "%b %d")+
  #   labs(x="", y="Expanded adult Chum", fill="Segment", colour="Segment") +
  #   theme_bw() +
  #   theme(axis.text.x = element_text(angle=45, hjust=1)),
  
  ggplot() +
    geom_bar(data=nit.sil %>% filter(species=="Coho") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
               summarize(Spawning=sum(`adults # spawning`),
                         Holding = sum(`adults # holding`)) %>% pivot_longer(cols=c(Spawning,Holding), names_to="behaviour", values_to="count"), 
             aes(x=as.Date(date), y=count, fill=behaviour, colour=`upper/lower (for Nitinat River only)`), 
             stat="identity", position="stack", alpha=0.7, size=1) +
    # geom_bar(data=nrh.cm,
    #          aes(x=as.Date(`Start Date`), y=`Total Count`*5), fill="black", stat="identity", alpha=0.7) +
    scale_y_continuous(sec.axis = sec_axis(~./5, name="Swim ins")) +
    scale_x_date(date_breaks="7 day", date_labels = "%b %d") +
    scale_fill_manual(breaks=waiver(), values=c("transparent", "gray")) +
    labs(x="", y="Spawning adult Coho", fill="Behaviour", colour="Behaviour") +
    theme_bw() +
    theme(axis.text.x = element_text(angle=45, hjust=1))
```

Figure: Expanded adult Coho counts (large bars) and swim-ins (narrow black bars). Adult bar fill colours indicate behaviour (spawning=gray, holding=white), bar line colour indicates river segment (lower=pink, upper=blue). 

<br>

<br>

## Preliminary escapement estimate (AUC/PL+D) 

<br>

Table: Raw Coho counts

```{r eval=F,include=F}
ver23 %>% 
  filter(ShortSpecies=="CO") %>% 
  select(-c(FabricatedZero)) %>% 
  mutate(across(c(LiveAdults:ExpDeadJacks), ~round(., 0))) %>% 
  kbl(align="c") %>%
  kable_paper("hover", full_width=T, position = "center") %>%
  collapse_rows(columns=c(1:4), valign="middle", target=1)
```

<br>

```{r eval=F,include=F}
ggplot() +
  geom_line(data=AUC23%>%filter(species=="Coho"), aes(x=as.Date(Date), y=ExpAdultFishDays)) +
  geom_point(data=AUC23%>%filter(species=="Coho"), 
             aes(x=as.Date(Date), y=ExpAdultFishDays, fill=surveyMethod, colour=surveyMethod), shape=21, size=4) +
  scale_fill_manual(breaks=c("heli", "snorkel"), values=c("orange", "blue")) +
  scale_colour_manual(breaks=c("heli", "snorkel"), values=c("orange", "blue")) +
  labs(x="", y="Fish-days", fill="", colour="") +
  theme_bw()
```

Figure: Coho AUC curve. 

<br>

Table: AUC math (SIL+AUC database math extracted: SIL+AUC_database_output_2024-01-03 - AUC.xlsx)

```{r eval=F,include=F}
AUC23 %>% 
  filter(species=="Coho") %>% 
  relocate(surveyMethod, .before=Date) %>% 
  kbl(align="c") %>%
  kable_paper("hover", full_width=T, position = "center") %>%
  collapse_rows(columns=c(1:4), valign="middle", target=1) 
```

<br>

Table: Escapement estimates for each survey life value.
```{r eval=F,include=F}
AUC23 %>% 
  filter(species=="Coho") %>%
  select(-c(species)) %>%
  summarize(fishDays = sum(ExpAdultFishDays),
            est_SL25 = round(fishDays/25,0),
            est_SL30 = round(fishDays/30,0),
            est_SL35 = round(fishDays/35,0)) %>% 
  pivot_longer(cols=c(est_SL25:est_SL35), names_to = "Rank", values_to="Estimate") %>% 
  mutate(Rank = case_when(grepl("25", Rank) ~ "Low Value",
                           grepl("30", Rank) ~ "Middle Value",
                           grepl("35", Rank) ~ "High Value"),
         SL = case_when(grepl("Low", Rank) ~ 25,
                        grepl("Middle", Rank) ~ 30,
                        grepl("High", Rank) ~ 35)) %>%
  relocate(SL, .after = Rank) %>%
  kbl(align="c") %>%
  kable_paper("hover", full_width=T, position = "center") %>%
  collapse_rows(columns=c(1:4), valign="middle", target=1) 
```
<br>

### Estimate decision:

For 2023 estimate we went with PL+D as even the highest AUC estimate didn't cover the peak raw live count. Expanded PL+D = ~1376 adults and ~156 jacks. 

<br>

<br>

## Tributaries

Live Coho were present in Campus, NoName and Worthless Creeks in 2023.

```{r eval=F,include=F}
tribs.sil %>% 
  filter(species=="Coho") %>% 
  group_by(system) %>% 
  summarize(exp_ad_live = sum(`adults live / habitat seen (# Est Live)`),
            dead_adults = sum(`adults dead`)) %>%
  mutate(across(c(exp_ad_live:dead_adults), ~round(., 0))) %>%
  kbl(align="c") %>%
  kable_paper("hover", full_width=T, position = "center") %>%
  collapse_rows(columns=c(1), valign="middle", target=1) 
```

<br>

<br>

## `r analysis_year` compared to historical run timing and abundance

```{r}
ggplot() +
  geom_point(data=nit.rtStWk%>%filter(species=="coho"),
           aes(x=stat_week, y=count, colour=plot_group, fill=plot_group, size=plot_group),
           shape=21, alpha=0.7) +
  scale_fill_manual(breaks=waiver(), values=c("gray60", "dodger blue")) +
  scale_colour_manual(breaks=waiver(), values=c("gray60", "dodger blue")) +
  scale_size_manual(breaks=waiver(), values=c(1.5,3.5)) +
  labs(x="Stat week", y="Live adult Coho", fill="Year:", colour="Year:", size="Year:") +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.box = "horizontal")
```
<br>

```{r}
ggplot()  +
  geom_bar(data=nit.esc %>% filter(`Waterbody Name`=="NITINAT RIVER", `Species Qualified`=="CO"), 
           aes(x=`Analysis Year`, y=`Natural Adult Spawners`), stat="identity", colour="dodger blue", fill="dodger blue", alpha=0.8) +
  geom_bar(data=nit.esc %>% filter(`Waterbody Name`=="NITINAT RIVER", `Species Qualified`=="CO"), 
           aes(x=`Analysis Year`, y=`Unspecified Return`), stat="identity", colour="gray80", fill="gray80", alpha=0.8) +  
  geom_bar(aes(x=analysis_year, y=1376), stat="identity", colour="orange", fill="orange", alpha=0.8) +   
  scale_x_continuous(breaks=seq(min(nit.esc$`Analysis Year`), analysis_year, by=2)) +
  labs(x="Return Year", y="Coho spawners") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

Figure: 2023 preliminary adult Coho escapement estimate (orange) relative to historical adult (blue) and unspecified (gray) escapement estimates. Gaps indicate years without estimates. 

<br>

<br>

<br>



























# **Sockeye**            
 
<br>

Snorkel and heli surveys were primarily timed for Chinook and Chum. Sockeye counts are incidental and should be considered low-quality/incomplete. Very low water this summer likely resulted in a barrier to migration, preventing sockeye from reaching Parker Creek until rains came much later. For this reason, SK counts in 2023 may be much higher than previous years as they were trapped in the lower mainstem and more visible to surveyors in late August.

<br>

## Raw vs expanded counts

Counts are expanded for both % habitat seen and observer efficiency, depending on survey conditions. 

```{r}
ggplot() +
  geom_bar(data=ver23 %>% filter(ShortSpecies=="SK") %>% pivot_longer(LiveAdults:ExpDeadJacks, names_to="est_type", values_to="n") %>% filter(est_type %in% c("LiveAdults", "ExpLiveAdults")), 
           aes(x=as.Date(Date), y=n, colour=est_type, fill=est_type), stat="identity", position="dodge", alpha=0.5) +
  labs(x="", y="Adult Sockeye counts", fill="Expansion", colour="Expansion") +
  scale_x_date(date_breaks="7 day", date_labels = "%b %d") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

Figure: Live Adult sockeye counts already expanded for % habitat seen (blue), and with additional % population expansion (pink). 

<br>

<br>

## Counts & environmentals

```{r}
ggplot() +
  geom_ribbon(data=nit.water %>% group_by(date) %>% summarize(min=min(gauge_m), max=max(gauge_m)), 
              aes(x=as.Date(date), ymin=min*100, ymax=max*100), fill="dodger blue", alpha=0.5) +
  geom_line(data=nit.water %>% group_by(date) %>% summarize(mean = mean(gauge_m)), 
            aes(x=as.Date(date), y=mean*100), colour="dodger blue") +
  
  geom_bar(data=nit.sil %>% filter(species=="Sockeye") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
             summarize(n=sum(`adults live / habitat seen (# Est Live)`)), 
           aes(x=as.Date(date), y=n, group=type, fill=`upper/lower (for Nitinat River only)`, colour=`upper/lower (for Nitinat River only)`), 
           stat="identity", position="stack", alpha=0.7) +

  scale_y_continuous(sec.axis = sec_axis(~./100, name="Staff gauge (m)")) +
  scale_x_date(date_breaks="7 day", date_labels="%b %d") +
  labs(x="", y="Expanded adult Sockeye", fill="Segment", colour="Segment") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

Figure: Sockeye counts (expanded adults) and water levels (blue line) 

<br>

<br>

## Behaviour & carcasses

```{r}
  ggplot() +
    geom_bar(data=nit.sil %>% filter(species=="Sockeye") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
               summarize(Spawning=sum(`adults # spawning`),
                         Holding = sum(`adults # holding`)) %>% pivot_longer(cols=c(Spawning,Holding), names_to="behaviour", values_to="count"), 
             aes(x=as.Date(date), y=count, fill=behaviour, colour=`upper/lower (for Nitinat River only)`), 
             stat="identity", position="stack", alpha=0.7, size=1) +
    geom_point(data=nit.sil %>% filter(species=="Sockeye") %>% group_by(date) %>% 
                 summarize(n=sum(`adults dead`)), 
               aes(x=as.Date(date), y=n), size=2, alpha=0.5) +
    geom_line(data=nit.sil %>% filter(species=="Sockeye") %>% group_by(date) %>% 
                summarize(n=sum(`adults dead`)), 
              aes(x=as.Date(date), y=n), 
              stat="identity", position="stack") +
    #scale_y_continuous(sec.axis = sec_axis(~./2, name="Carcasses")) +
    scale_x_date(date_breaks="7 day", date_labels = "%b %d")+
    scale_fill_manual(breaks=waiver(), values=c("transparent", "gray")) +
    labs(x="", y="Spawning adult Sockeye", fill="Behaviour", colour="Segment") +
    theme_bw() +
    theme(axis.text.x = element_text(angle=45, hjust=1))
```
<br>

Figure: Top: Expanded live Sockeye (bars) and carcasses (black line). Bottom: Spawning (gray bars) and holding (white bars) adult Sockeye and carcasses (black line). 

<br>

<br>

## Snorkel & swim-in counts

```{r eval=F, include=F}
# Figure: Top: Chum counts (expanded adults) and swim-ins (gray bars). Bottom: Spawning (gray) and holding (white) Chum and swim-ins (black bars).
```

There are no swim-ins for Nitinat sockeye as they are not enhanced. 

<br>

<br>

## Preliminary escapement estimate (AUC/PL+D) 

<br>

Table: Raw Sockeye counts

```{r}
ver23 %>% 
  filter(ShortSpecies=="SK") %>% 
  mutate(across(c(LiveAdults:ExpDeadJacks), ~round(., 0))) %>% 
  kbl(align="c") %>%
  kable_paper("hover", full_width=T, position = "center") %>%
  collapse_rows(columns=c(1:4), valign="middle", target=1) 
```

<br>

```{r}
ggplot() +
  geom_line(data=AUC23%>%filter(species=="Sockeye"), aes(x=as.Date(Date), y=ExpAdultFishDays)) +
  geom_point(data=AUC23%>%filter(species=="Sockeye"), aes(x=as.Date(Date), y=ExpAdultFishDays), shape=21, size=4, fill="blue") +
  labs(x="", y="Fish-days", fill="", colour="") +
  theme_bw()
```

Figure: AUC curve for Sockeye

<br>

Table: AUC math (SIL+AUC database math extracted: SIL+AUC_database_output_2024-01-03 - AUC.xlsx)

```{r}
AUC23 %>% 
  filter(species=="Sockeye") %>% 
  kbl(align="c") %>%
  kable_paper("hover", full_width=T, position = "center") %>%
  collapse_rows(columns=c(1:4), valign="middle", target=1) 
```

<br>

<br>

### Estimate decision:

For 2023 estimate we went with expanded PL+D. Low water was a migration barrier and likely meant sockeye were much more visible to surveyors later in the season than usual (they were prevented from migrating further upstream to Parker Creek); 2023 estimate may be much higher than previous years. The Expanded PL+D adult estimate for 2023 = ~460; expanded jacks = ~40.  *This is subject to change; see EscBatchHardCopy report for specific estimate abundance.*

<br>

<br>

## Tributaries

Sockeye were present in Hobiton and Campus Creeks in 2023, but the primary enumeration method is by NTC and DFN echosounder. 
Hobiton was swum once late May due to delays in the echosounder installation, and again in late August during drought conditions to assess whether pre-spawn mortalities were occurring in the creek. Only one pre-spawn jack was noted, the remaining dead were too deep or too decomposed to assess mortality. 

```{r}
tribs.sil %>% 
  filter(species=="Sockeye") %>% 
  group_by(date, system) %>% 
  summarize(exp_ad_live = sum(`adults live / habitat seen (# Est Live)`),
            dead_adults = sum(`adults dead`),
            jk_live = sum(`jacks live`),
            jk_dead = sum(`jacks dead`)) %>%
  mutate(across(c(exp_ad_live:dead_adults), ~round(., 0))) %>%
  kbl(align="c") %>%
  kable_paper("hover", full_width=T, position = "center") %>%
  collapse_rows(columns=c(1), valign="middle", target=1) 
```

<br>

<br>

## `r analysis_year` compared to historical run timing and abundance

Run timing information is not available for Nitinat Sockeye.

<br>

```{r}
ggplot()  +
  geom_bar(data=nit.esc %>% filter(`Waterbody Name`=="NITINAT RIVER", `Species Qualified`=="SEL"), 
           aes(x=`Analysis Year`, y=`Natural Adult Spawners`), stat="identity", colour="dodger blue", fill="dodger blue", alpha=0.8) +
  geom_bar(data=nit.esc %>% filter(`Waterbody Name`=="NITINAT RIVER", `Species Qualified`=="SEL"), 
           aes(x=`Analysis Year`, y=`Unspecified Return`), stat="identity", colour="gray80", fill="gray80", alpha=0.8) +  
  geom_bar(aes(x=analysis_year, y=421), stat="identity", colour="orange", fill="orange", alpha=0.8) +   
  scale_x_continuous(breaks=seq(min(nit.esc$`Analysis Year`), analysis_year, by=2)) +
  labs(x="Return Year", y="Sockeye spawners") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

Figure: 2023 preliminary adult Sockeye escapement estimate (orange) relative to historical adult (blue) and unspecified (gray) escapement estimates. Gaps indicate years without estimates. 

<br>

<br>
