---
title: "Nitinat escapement 2023 (PRELIM)"
author: "South Coast StA"
date: "Last update: `r Sys.Date()`"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_collapsed: false 
    toc_depth: 5
    number_sections: true
knit: (function(inputFile, encoding) { 
  rmarkdown::render(inputFile, encoding=encoding, output_file=paste0(substr(inputFile,1,nchar(inputFile)-4), "_PRELIMINARY_", format(Sys.Date(), sep="_", "%Y-%m-%d"), '.html')) 
  }
  )
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning=F, message=F)

# Load libraries ----------------------------
library(tidyverse)
library(readxl)
library(here)
library(ggpubr)
library(kableExtra)


# ============================= Load data ============================

# SILS ----------------------------
nit.sil <- readxl::read_excel("//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/WCVI/ESCAPEMENT/Data/2023/SILs/Area 21-22/Nitinat River/A22_Nitinat_River_2023  kd.xlsx", sheet="Nitinat River raw SIL entry", guess_max=20000)

auc23 <- readxl::read_excel("//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/WCVI/ESCAPEMENT/Data/2023/SILs/Area 21-22/Nitinat River/A22_Nitinat_River_2023  kd.xlsx", sheet="AUC_R", guess_max=20000)

tribs.sil <- readxl::read_excel("//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/WCVI/ESCAPEMENT/Data/2023/SILs/Area 21-22/Tribs/A22_Tributaries_2023.xlsx", sheet="Tributaries Raw SIL entry", guess_max=20000)


# Hatchery ----------------------------
nrh.cn <- readxl::read_excel("//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/WCVI/ESCAPEMENT/Data/2023/SILs/Area 21-22/Adult_Management - Nitinat CN 2023.xlsx", sheet="Adult Management", guess_max=20000, skip=1)

nrh.cm <- readxl::read_excel("//dcbcpbsna01a.ENT.dfo-mpo.ca/SCD_Stad/WCVI/ESCAPEMENT/Data/2023/SILs/Area 21-22/Nitinat Chum Removals_By Date_Nov 29_2023.xlsx", sheet="All Adult Activities", skip=1) 


# Enviro ----------------------------
nit.water <- list.files(here::here("data", "enviro"), pattern = "NitinatWaterLevelHydrometData", full.names = T) %>% 
  map(~readxl::read_excel(path=.x, sheet="Exported Data", guess_max=20000)) %>%
  purrr::list_rbind() %>%
  mutate(gauge_m = case_when(grepl("Estimated", `Staff Gauge(mH20)`) ~ NA,
                             TRUE ~ as.numeric(`Staff Gauge(mH20)`)),
         sensor_depth = case_when(grepl("Estimated", `Sensor Depth(mH20)`) ~ NA,
                             TRUE ~ as.numeric(`Sensor Depth(mH20)`)),
         date = as.Date(Time)) %>% 
  filter(date<= as.Date("2023-11-08"))
```

<br>

<br>


# **Chinook**

Six snorkel surveys and one helicopter survey were conducted for Chinook in 2023. 

<br>

## Counts & environmentals

```{r }
ggplot() +
  geom_ribbon(data=nit.water %>% group_by(date) %>% summarize(min=min(gauge_m), max=max(gauge_m)), 
              aes(x=as.Date(date), ymin=min*5000, ymax=max*5000), fill="dodger blue", alpha=0.5) +
  geom_line(data=nit.water %>% group_by(date) %>% summarize(mean = mean(gauge_m)), 
            aes(x=as.Date(date), y=mean*5000), colour="dodger blue") +
  geom_bar(data=nit.sil %>% filter(species=="Chinook") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
             summarize(n=sum(`adults live / habitat seen (# Est Live)`)), 
           aes(x=as.Date(date), y=n, group=type, fill=`upper/lower (for Nitinat River only)`, colour=`upper/lower (for Nitinat River only)`), 
           stat="identity", position="stack", alpha=0.7) +
  # geom_line(data=nit.sil %>% filter(species=="Chinook") %>% group_by(date) %>% 
  #             summarize(n=sum(`adults dead`)), 
  #           aes(x=as.Date(date), y=n), 
  #           stat="identity", position="stack") +
  labs(x="", y="Expanded adult Chinook") +
  scale_y_continuous(sec.axis = sec_axis(~./5000, name="Staff gauge (m)")) +
  scale_x_date(date_breaks="7 day", date_labels="%b %d") +
  labs(x="", Y="Live chinook", fill="Segment", colour="Segment") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

Figure: Chinook counts (expanded adults) and water levels (blue line) 

<br>

<br>

## Behaviour & carcasses

```{r}
ggarrange(
  ggplot() +
    geom_bar(data=nit.sil %>% filter(species=="Chinook") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
               summarize(n=sum(`adults live / habitat seen (# Est Live)`)), 
             aes(x=as.Date(date), y=n, fill=`upper/lower (for Nitinat River only)`, colour=`upper/lower (for Nitinat River only)`), 
             stat="identity", position="stack", alpha=0.7) +
    geom_point(data=nit.sil %>% filter(species=="Chinook") %>% group_by(date) %>% 
                 summarize(n=sum(`adults dead`)), 
               aes(x=as.Date(date), y=n*10), size=2, alpha=0.5) +
    geom_line(data=nit.sil %>% filter(species=="Chinook") %>% group_by(date) %>% 
                summarize(n=sum(`adults dead`)), 
              aes(x=as.Date(date), y=n*10), 
              stat="identity", position="stack") +
    scale_y_continuous(sec.axis = sec_axis(~./10, name="Carcasses")) +
    scale_x_date(date_breaks="7 day", date_labels = "%b %d")+
    labs(x="", y="Expanded live adult Chinook", fill="Segment", colour="Segment") +
    theme_bw() +
    theme(axis.text.x = element_text(angle=45, hjust=1)),
  
  ggplot() +
    geom_bar(data=nit.sil %>% filter(species=="Chinook") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
               summarize(Spawning=sum(`adults # spawning`),
                         Holding = sum(`adults # holding`)) %>% pivot_longer(cols=c(Spawning,Holding), names_to="behaviour", values_to="count"), 
             aes(x=as.Date(date), y=count, fill=behaviour, colour=`upper/lower (for Nitinat River only)`), 
             stat="identity", position="stack", alpha=0.7, size=1) +
    geom_point(data=nit.sil %>% filter(species=="Chinook") %>% group_by(date) %>% 
                 summarize(n=sum(`adults dead`)), 
               aes(x=as.Date(date), y=n*10), size=2, alpha=0.5) +
    geom_line(data=nit.sil %>% filter(species=="Chinook") %>% group_by(date) %>% 
                summarize(n=sum(`adults dead`)), 
              aes(x=as.Date(date), y=n*10), 
              stat="identity", position="stack") +
    scale_y_continuous(sec.axis = sec_axis(~./10, name="Carcasses")) +
    scale_x_date(date_breaks="7 day", date_labels = "%b %d")+
    scale_fill_manual(breaks=waiver(), values=c("transparent", "gray")) +
    labs(x="", y="Spawning adult Chinook", fill="Behaviour", colour="Segment") +
    theme_bw() +
    theme(axis.text.x = element_text(angle=45, hjust=1)),
  
  nrow=2 
)
```
<br>

Figure: Top: Expanded live Chinook (bars) and carcasses (black line). Bottom: Spawning (gray bars) and holding (white bars) adult Chinook and carcasses (black line). 

<br>

<br>

## Snorkel & swim-in counts

```{r}
ggarrange(
  ggplot() +
    geom_bar(data=nit.sil %>% filter(species=="Chinook") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
               summarize(n=sum(`adults live / habitat seen (# Est Live)`)), 
             aes(x=as.Date(date), y=n, group=type, fill=`upper/lower (for Nitinat River only)`, colour=`upper/lower (for Nitinat River only)`), 
             stat="identity", position="stack", alpha=0.7) + 
    geom_bar(data=nrh.cn %>% filter(`Parent Activity Type`=="Adult Capture", grepl("Swim Ins", `Inventory Lot`)) %>% 
               group_by(`Activity Start Date`) %>% summarize(n=sum(`Total Count`)),
             aes(x=as.Date(`Activity Start Date`), y=n*1), fill="black", stat="identity", alpha=0.7) +
    scale_y_continuous(sec.axis = sec_axis(~./1, name="Swim ins")) +
    scale_x_date(date_breaks="7 day", date_labels = "%b %d")+
    labs(x="", y="Expanded adult Chinook", fill="Segment", colour="Segment") +
    theme_bw() +
    theme(axis.text.x = element_text(angle=45, hjust=1)),
  
  ggplot() +
    geom_bar(data=nit.sil %>% filter(species=="Chinook") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
               summarize(Spawning=sum(`adults # spawning`),
                         Holding = sum(`adults # holding`)) %>% pivot_longer(cols=c(Spawning,Holding), names_to="behaviour", values_to="count"), 
             aes(x=as.Date(date), y=count, fill=behaviour, colour=`upper/lower (for Nitinat River only)`), 
             stat="identity", position="stack", alpha=0.7, size=1) +
    geom_bar(data=nrh.cn %>% filter(`Parent Activity Type`=="Adult Capture", grepl("Swim Ins", `Inventory Lot`)) %>% 
               group_by(`Activity Start Date`) %>% summarize(n=sum(`Total Count`)),
             aes(x=as.Date(`Activity Start Date`), y=n*1), fill="black", stat="identity", alpha=0.7) +
    scale_y_continuous(sec.axis = sec_axis(~./1, name="Swim ins")) +
    scale_x_date(date_breaks="7 day", date_labels = "%b %d") +
    scale_fill_manual(breaks=waiver(), values=c("transparent", "gray")) +
    labs(x="", y="Spawning adult Chinook", fill="Behaviour", colour="Behaviour") +
    theme_bw() +
    theme(axis.text.x = element_text(angle=45, hjust=1)),
  
  nrow=2 
)
```

Figure: Top: Chinook counts (expanded adults) and swim-ins (gray bars). Bottom: Spawning (gray) and holding (white) Chinook and swim-ins (black bars). 

<br>

<br>

## Preliminary escapement estimate (AUC) 

<br>

```{r}
ggarrange(
ggplot() +
  geom_line(data=auc23%>%filter(method=="with heli", species=="chinook"), aes(x=as.Date(survey_date), y=`expanded adult`)) +
  geom_point(data=auc23%>%filter(method=="with heli", species=="chinook"), aes(x=as.Date(survey_date), y=`expanded adult`, fill=type, colour=type), shape=21, size=4) +
  scale_fill_manual(breaks=c("heli", "snorkel"), values=c("orange", "blue")) +
  scale_colour_manual(breaks=c("heli", "snorkel"), values=c("orange", "blue")) +
  labs(x="", y="Fish-days", fill="", colour="") +
  theme_bw(),

ggplot() +
  geom_line(data=auc23%>%filter(method=="without heli", species=="chinook"), aes(x=as.Date(survey_date), y=`expanded adult`)) +
  geom_point(data=auc23%>%filter(method=="without heli", species=="chinook"), aes(x=as.Date(survey_date), y=`expanded adult`, fill=type, colour=type), shape=21, size=4) +
  scale_fill_manual(breaks=c("heli", "snorkel"), values=c("orange", "blue")) +
  scale_colour_manual(breaks=c("heli", "snorkel"), values=c("orange", "blue")) +
  labs(x="", y="Fish-days", fill="", colour="") +
  theme_bw(),
nrow=2,
common.legend = T, legend = "bottom"
)
```

Figure: AUC curves with heli (top) and without heli (bottom) survey. 

<br>

Table: Raw chinook counts

```{r}
nit.sil %>% 
  filter(species=="Chinook") %>% 
  group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
  summarize(adults_live = sum(`adults live`),
            adults_spawning = sum(`adults # spawning`),
            adults_holding = sum(`adults # holding`),
            dead_adults = sum(`adults dead`),
            EXPANDED_LIVE_ADULTS = sum(`adults live / habitat seen (# Est Live)`),
            
            jacks_live = sum(`jacks live`),
            jacks_spawning = sum(`jacks # spawning`),
            jacks_holding = sum(`jacks # holding`),
            dead_jacks = sum(`jacks dead`)) %>% 
  mutate(across(c(adults_live:dead_jacks), ~round(., 0))) %>% 
  rename(section=`upper/lower (for Nitinat River only)`) %>% 
  kbl(align="c") %>%
  kable_paper("hover", full_width=T, position = "center") %>%
  collapse_rows(columns=c(1:4), valign="middle", target=1) 
```


<br>

Table: AUC math (copy of Z:/WCVI/ESCAPEMENT/Data/2023/SILs/Area 21-22/Nitinat River/A22_Nitinat_River_2023)

```{r}
auc23 %>% 
  filter(species=="chinook") %>% 
  select(-c(species)) %>%
  relocate(method, .before=survey_date) %>% 
  kbl(align="c") %>%
  kable_paper("hover", full_width=T, position = "center") %>%
  collapse_rows(columns=c(1:4), valign="middle", target=1) 
```

<br>

Table: Escapement estimates with/without helicopter survey for each survey life value.
```{r}
auc23 %>% 
  filter(species=="chinook") %>%
  select(-c(species)) %>%
  group_by(method) %>% 
  summarize(fish_days_SL15 = round(sum(`expanded adult`)/15,0),
            fish_days_SL20 = round(sum(`expanded adult`)/20,0),
            fish_days_SL25 = round(sum(`expanded adult`)/25,0)) %>% 
  pivot_longer(cols=c(fish_days_SL15:fish_days_SL25), names_to = "Rank", values_to="Estimate") %>% 
  mutate(Rank = case_when(grepl("15", Rank) ~ "Low Value",
                           grepl("20", Rank) ~ "Middle Value",
                           grepl("25", Rank) ~ "High Value"),
         SL = case_when(grepl("Low", Rank) ~ 15,
                        grepl("Middle", Rank) ~ 20,
                        grepl("High", Rank) ~ 25),
         Est_without_swimins = Estimate-6846) %>%
  relocate(SL, .after = Rank) %>%
  kbl(align="c") %>%
  kable_paper("hover", full_width=T, position = "center") %>%
  collapse_rows(columns=c(1:4), valign="middle", target=1) 
```
<br>

For 2023 estimate we went with SL=15 for three reasons: First, there were a lot more swim-ins than usual this year, which was attributed to the COVID releases leaving the hatchery directly rather than being reared in the lake pen. Fish homing to the hatchery may have had more directed migration and therefore not stayed in the system long. Second, the higher estimate from the shorter SL resulted in an estimate that was high enough to cover the raw peak live count of ~15k,  but once removing the 6846 swim-ins still resulted in a reasonable estimate that agreed with what the hatchery observed. Third, there may have been a slight delay in entry as the water was low for the early season and rains were a bit later than usual (though not as late as 2022). 

<br>

<br>

## Tributaries

Chinook were not seen in any Nitinat River or Lake tributaries in 2023.

```{r}
tribs.sil %>% 
  filter(species=="Chinook") %>% 
  group_by(system) %>% 
  summarize(exp_ad_live = sum(`adults live / habitat seen (# Est Live)`))
```
<br>

<br>

<br>













# Chum 

<br>

Six snorkel surveys and one helicopter survey were conducted for Chum in 2023. The heli survey may have been slightly early for Chum. 

<br>

## Counts & environmentals

```{r }
ggplot() +
  geom_ribbon(data=nit.water %>% group_by(date) %>% summarize(min=min(gauge_m), max=max(gauge_m)), 
              aes(x=as.Date(date), ymin=min*5000, ymax=max*5000), fill="dodger blue", alpha=0.5) +
  geom_line(data=nit.water %>% group_by(date) %>% summarize(mean = mean(gauge_m)), 
            aes(x=as.Date(date), y=mean*5000), colour="dodger blue") +
  geom_bar(data=nit.sil %>% filter(species=="Chum") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
             summarize(n=sum(`adults live / habitat seen (# Est Live)`)), 
           aes(x=as.Date(date), y=n, group=type, fill=`upper/lower (for Nitinat River only)`, colour=`upper/lower (for Nitinat River only)`), 
           stat="identity", position="stack", alpha=0.7) +
  # geom_line(data=nit.sil %>% filter(species=="Chum") %>% group_by(date) %>% 
  #             summarize(n=sum(`adults dead`)), 
  #           aes(x=as.Date(date), y=n), 
  #           stat="identity", position="stack") +
  scale_y_continuous(sec.axis = sec_axis(~./5000, name="Staff gauge (m)")) +
  scale_x_date(date_breaks="7 day", date_labels="%b %d") +
  labs(x="", y="Expanded adult Chum", fill="Segment", colour="Segment") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

Figure: Chinook counts (expanded adults) and water levels (blue line) 

<br>

<br>

## Behaviour & carcasses

```{r}
ggarrange(
  ggplot() +
    geom_bar(data=nit.sil %>% filter(species=="Chum") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
               summarize(n=sum(`adults live / habitat seen (# Est Live)`)), 
             aes(x=as.Date(date), y=n, fill=`upper/lower (for Nitinat River only)`, colour=`upper/lower (for Nitinat River only)`), 
             stat="identity", position="stack", alpha=0.7) +
    geom_point(data=nit.sil %>% filter(species=="Chum") %>% group_by(date) %>% 
                 summarize(n=sum(`adults dead`)), 
               aes(x=as.Date(date), y=n*10), size=2, alpha=0.5) +
    geom_line(data=nit.sil %>% filter(species=="Chum") %>% group_by(date) %>% 
                summarize(n=sum(`adults dead`)), 
              aes(x=as.Date(date), y=n*10), 
              stat="identity", position="stack") +
    scale_y_continuous(sec.axis = sec_axis(~./10, name="Carcasses")) +
    scale_x_date(date_breaks="7 day", date_labels = "%b %d")+
    labs(x="", y="Expanded live adult Chum", fill="Segment", colour="Segment") +
    theme_bw() +
    theme(axis.text.x = element_text(angle=45, hjust=1)),
  
  ggplot() +
    geom_bar(data=nit.sil %>% filter(species=="Chum") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
               summarize(Spawning=sum(`adults # spawning`),
                         Holding = sum(`adults # holding`)) %>% pivot_longer(cols=c(Spawning,Holding), names_to="behaviour", values_to="count"), 
             aes(x=as.Date(date), y=count, fill=behaviour, colour=`upper/lower (for Nitinat River only)`), 
             stat="identity", position="stack", alpha=0.7, size=1) +
    geom_point(data=nit.sil %>% filter(species=="Chum") %>% group_by(date) %>% 
                 summarize(n=sum(`adults dead`)), 
               aes(x=as.Date(date), y=n*10), size=2, alpha=0.5) +
    geom_line(data=nit.sil %>% filter(species=="Chum") %>% group_by(date) %>% 
                summarize(n=sum(`adults dead`)), 
              aes(x=as.Date(date), y=n*10), 
              stat="identity", position="stack") +
    scale_y_continuous(sec.axis = sec_axis(~./10, name="Carcasses")) +
    scale_x_date(date_breaks="7 day", date_labels = "%b %d")+
    scale_fill_manual(breaks=waiver(), values=c("transparent", "gray")) +
    labs(x="", y="Spawning adult Chum", fill="Behaviour", colour="Segment") +
    theme_bw() +
    theme(axis.text.x = element_text(angle=45, hjust=1)),
  
  nrow=2 
)
```
<br>

Figure: Top: Expanded live Chum (bars) and carcasses (black line). Bottom: Spawning (gray bars) and holding (white bars) adult Chum and carcasses (black line). 

<br>

<br>

## Snorkel & swim-in counts

```{r}
ggarrange(
  ggplot() +
    geom_bar(data=nit.sil %>% filter(species=="Chum") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
               summarize(n=sum(`adults live / habitat seen (# Est Live)`)), 
             aes(x=as.Date(date), y=n, group=type, fill=`upper/lower (for Nitinat River only)`, colour=`upper/lower (for Nitinat River only)`), 
             stat="identity", position="stack", alpha=0.7) + 
    geom_bar(data=nrh.cm, 
             aes(x=as.Date(`Start Date`), y=`Total Count`*5), fill="black", stat="identity", alpha=0.7) +
    scale_y_continuous(sec.axis = sec_axis(~./5, name="Swim ins")) +
    scale_x_date(date_breaks="7 day", date_labels = "%b %d")+
    labs(x="", y="Expanded adult Chum", fill="Segment", colour="Segment") +
    theme_bw() +
    theme(axis.text.x = element_text(angle=45, hjust=1)),
  
  ggplot() +
    geom_bar(data=nit.sil %>% filter(species=="Chum") %>% group_by(date, type, `upper/lower (for Nitinat River only)`) %>% 
               summarize(Spawning=sum(`adults # spawning`),
                         Holding = sum(`adults # holding`)) %>% pivot_longer(cols=c(Spawning,Holding), names_to="behaviour", values_to="count"), 
             aes(x=as.Date(date), y=count, fill=behaviour, colour=`upper/lower (for Nitinat River only)`), 
             stat="identity", position="stack", alpha=0.7, size=1) +
    geom_bar(data=nrh.cm,
             aes(x=as.Date(`Start Date`), y=`Total Count`*5), fill="black", stat="identity", alpha=0.7) +
    scale_y_continuous(sec.axis = sec_axis(~./5, name="Swim ins")) +
    scale_x_date(date_breaks="7 day", date_labels = "%b %d") +
    scale_fill_manual(breaks=waiver(), values=c("transparent", "gray")) +
    labs(x="", y="Spawning adult Chum", fill="Behaviour", colour="Behaviour") +
    theme_bw() +
    theme(axis.text.x = element_text(angle=45, hjust=1)),
  
  nrow=2 
)
```

Figure: Top: Chum counts (expanded adults) and swim-ins (gray bars). Bottom: Spawning (gray) and holding (white) Chum and swim-ins (black bars). 

<br>

<br>

## Preliminary escapement estimate (AUC) 

<br>

```{r}
ggarrange(
ggplot() +
  geom_line(data=auc23%>%filter(method=="with heli"), aes(x=as.Date(survey_date), y=`expanded adult`)) +
  geom_point(data=auc23%>%filter(method=="with heli"), aes(x=as.Date(survey_date), y=`expanded adult`, fill=type, colour=type), shape=21, size=4) +
  scale_fill_manual(breaks=c("heli", "snorkel"), values=c("orange", "blue")) +
  scale_colour_manual(breaks=c("heli", "snorkel"), values=c("orange", "blue")) +
  labs(x="", y="Fish-days", fill="", colour="") +
  theme_bw(),

ggplot() +
  geom_line(data=auc23%>%filter(method=="without heli"), aes(x=as.Date(survey_date), y=`expanded adult`)) +
  geom_point(data=auc23%>%filter(method=="without heli"), aes(x=as.Date(survey_date), y=`expanded adult`, fill=type, colour=type), shape=21, size=4) +
  scale_fill_manual(breaks=c("heli", "snorkel"), values=c("orange", "blue")) +
  scale_colour_manual(breaks=c("heli", "snorkel"), values=c("orange", "blue")) +
  labs(x="", y="Fish-days", fill="", colour="") +
  theme_bw(),
nrow=2,
common.legend = T, legend = "bottom"
)
```

Figure: AUC curves with heli (top) and without heli (bottom) survey. 

<br>

Table: AUC math (copy of Z:/WCVI/ESCAPEMENT/Data/2023/SILs/Area 21-22/Nitinat River/A22_Nitinat_River_2023)

```{r}
auc23 %>% 
  relocate(method, .before=survey_date) %>% 
  kbl(align="c") %>%
  kable_paper("hover", full_width=T, position = "center") %>%
  collapse_rows(columns=c(1:4), valign="middle", target=1) 
```

<br>

Table: Escapement estimates with/without helicopter survey for each survey life value.
```{r}
auc23 %>% 
  group_by(method) %>% 
  summarize(fish_days_SL15 = round(sum(`expanded adult`)/15,0),
            fish_days_SL20 = round(sum(`expanded adult`)/20,0),
            fish_days_SL25 = round(sum(`expanded adult`)/25,0)) %>% 
  pivot_longer(cols=c(fish_days_SL15:fish_days_SL25), names_to = "Rank", values_to="Estimate") %>% 
  mutate(Rank = case_when(grepl("15", Rank) ~ "Low Value",
                           grepl("20", Rank) ~ "Middle Value",
                           grepl("25", Rank) ~ "High Value"),
         SL = case_when(grepl("Low", Rank) ~ 15,
                        grepl("Middle", Rank) ~ 20,
                        grepl("High", Rank) ~ 25),
         Est_without_swimins = Estimate-6846) %>%
  relocate(SL, .after = Rank) %>%
  kbl(align="c") %>%
  kable_paper("hover", full_width=T, position = "center") %>%
  collapse_rows(columns=c(1:4), valign="middle", target=1) 
```
<br>

For 2023 estimate we went with SL=15 for three reasons: First, there were a lot more swim-ins than usual this year, which was attributed to the COVID releases leaving the hatchery directly rather than being reared in the lake pen. Fish homing to the hatchery may have had more directed migration and therefore not stayed in the system long. Second, the higher estimate from the shorter SL resulted in an estimate that was high enough to cover the raw peak live count of ~15k,  but once removing the 6846 swim-ins still resulted in a reasonable estimate that agreed with what the hatchery observed. Third, there may have been a slight delay in entry as the water was low for the early season and rains were a bit later than usual (though not as late as 2022). 

<br>

<br>

## Tributaries

Chinook were not seen in any Nitinat River or Lake tributaries in 2023.

```{r}
tribs.sil %>% 
  filter(species=="Chinook") %>% 
  group_by(system) %>% 
  summarize(exp_ad_live = sum(`adults live / habitat seen (# Est Live)`))
```
<br>

<br>

<br>




